---
title: "![](./logo.svg){height=40} \n Individual analysis results"
output:
  html_document: 
    fig_caption: yes
    theme: flatly
    css: custom.css
params:
  foldername:
    label: "Enter folder name"
    value: folder-name
    input: text
  seqname:
    label: "Enter sequence name"
    value: sequence-name
    input: text
  collection:
    label: "Enter collection name"
    value: collection
    input: text
  group:
    label: "Enter group name"
    value: group1
    input: text
  replicate:
    label: "Enter replicate number (direct-BSP)"
    value: 1
    input: numeric
  clone:
    label: "Enter clone number (cloning-BSP)"
    value: 0
    input: numeric
  DNA_seq:
    label: "Select .fasta file of reference DNA sequence from plus strand with header containing coordinates and strand (plus or minus) used for primer design"
    value: "DNA_seq.fasta"
    input: file
  genome:
    label: "Select genome"
    value: BSgenome.Hsapiens.UCSC.hg19
    input: select
    choices: !r installed.genomes()
    selected: BSgenome.Hsapiens.UCSC.hg19
  date_s1:
    label: "Select date of sequencing #1"
    value: !r Sys.Date()
    input: date
  date_s2:
    label: "Select date of sequencing #2"
    value: !r Sys.Date()
    input: date
  ab1_s1:
    label: "Select .ab1 file of sequencing result from one direction #1"
    value: "seq_s1.ab1"
    input: file
  ab1_s2:
    label: "Select .ab1 file of sequencing result from the other direction #2"
    value: "seq_s2.ab1"
    input: file
editor_options: 
  chunk_output_type: console
---

```{r setup, include=F}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_knit$set(root.dir = gsub("/scripts","",getwd()))
options(knitr.table.format = "html")
options(ucscChromosomeNames=FALSE)
Sys.setlocale(category = "LC_ALL", locale = "en_SE.utf8")
```

```{r Library sync, include=F}
# Restore package versions from the renv.lock file : synchronize library with lockfile
renv::restore(prompt=F)
```

```{r Manual parameters and files for TESTS, eval=F}
## TO TEST THE SCRIPT WITHOUT KNITTING

# Parameters variables
folder_name <- "Example data"
seq_name = "CDH1"
seq_coll = "" ; seq_coll = ifelse(seq_coll==0,"",seq_coll)
seq_group = "High methylated DNA"
seq_rep = "3"  ; seq_rep = ifelse(seq_rep==0,"",seq_rep)
seq_clone = "0" ; seq_clone = ifelse(seq_clone=="0"|seq_clone==""|seq_clone==" ","",paste0("clone",str_pad(seq_clone, 2, "0", side="left")))
clone_nb = 0
seq_date_s1 = "2022-05-02"
seq_date_s2 = "2022-05-02"
selected_genome = "BSgenome.Hsapiens.UCSC.hg19"

# Get files paths using choose.files function (Windows only)
DNA_seq_path = choose.files(default = "", caption = "Select .fasta file of reference DNA sequence",
                                multi = FALSE, filters=Filters, index=nrow(Filters))
sangerseq_s1_path = choose.files(default = "", caption = "Select .ab1 file of sequencing result from one direction #1",
                                multi = FALSE, filters=Filters, index=nrow(Filters))
sangerseq_s2_path = choose.files(default = "", caption = "Select .ab1 file of sequencing result from the other direction #2",
                                multi = FALSE, filters=Filters, index=nrow(Filters))

useSeqs1 <- T
useSeqs2 <- T
```

```{r Read parameters from YAML header, include=F}
## SCRIPT TO USE PARAMS FOR KNITING : DO NOT RUN IF TESTING THE SCRIPT WITHOUT KNITTING

# Parameters variables
folder_name <- params$foldername
seq_name <- params$seqname
seq_coll <- params$collection ; seq_coll <- ifelse(seq_coll==0 | is.na(seq_coll),"",seq_coll)
seq_group <- params$group
if (params$replicate==0 | is.na(params$replicate)) {seq_rep <- ""} else {seq_rep <- params$replicate}
if (params$clone==0 | is.na(params$clone)) {seq_clone <- ""} else {seq_clone <- paste0("clone",str_pad(params$clone, 2, "0", side="left"))}
if (params$clone==0 | is.na(params$clone)) {clone_nb <- NA} else {clone_nb <- params$clone}

# Date 
if(length(params$date_s1)==0) {seq_date_s1 <- ""
} else if (is.null(params$date_s1) | is.na(params$date_s1)) { seq_date_s1 <- ""
} else {seq_date_s1 <- params$date_s1}

if(length(params$date_s2)==0) {seq_date_s2 <- ""
} else if (is.null(params$date_s2) | is.na(params$date_s2)) { seq_date_s2 <- ""
} else {seq_date_s2 <- params$date_s2}

# Date format
seq_date_s1 <- as.character(seq_date_s1)
seq_date_s2 <- as.character(seq_date_s2)

selected_genome <- params$genome

# Files

if(length(params$ab1_s1)==0) { useSeqs1 <- F
} else if (params$ab1_s1=="seq_s1.ab1" | is.na(params$ab1_s1) | is.null(params$ab1_s1) | params$ab1_s1=="") { useSeqs1 <- F
} else  { useSeqs1 <- T }

if(length(params$ab1_s2)==0) { useSeqs2 <- F
} else if (params$ab1_s2=="seq_s2.ab1" | is.na(params$ab1_s2) | is.null(params$ab1_s2) | params$ab1_s2=="") { useSeqs2 <- F
} else  { useSeqs2 <- T }

DNA_seq_path <- params$DNA_seq
sangerseq_s1_path <- params$ab1_s1
sangerseq_s2_path <- params$ab1_s2

```

```{r read files and extract coordinates, include=F, warning=T}
# SEQUENCE FASTA FILE

# Check fasta format
if (grepl("*.fasta$",DNA_seq_path)==F) { stop("sequence file must be in .fasta format") }

# Read fasta file as DNA String set
DNA_seq_stringset <- readDNAStringSet(DNA_seq_path,format="fasta")

# Set ref DNA sequence as a character variable
DNA_seq <- base::as.character(DNA_seq_stringset, use.names=F)

# Extract fasta header
DNA_seq_header <- DNA_seq_stringset@ranges@NAMES

```

```{r get information, include=F, warning=T}
# Extractcoordinates : chromosome
seq_chr <- stringr::str_extract(DNA_seq_header,"(?<=chr)(.{1,5})(?=:\\d{1,20}-\\d{1,20})")
seq_chr <- paste0("chr",seq_chr)
# Extract coordinates : start
seq_start <- as.numeric(stringr::str_extract(DNA_seq_header, "(?<=chr.{1,20}:)(\\d*)(?=-\\d{1,20})"))
# Extract coordinates : end
seq_end <- as.numeric(stringr::str_extract(DNA_seq_header, "(?<=chr.{1,20}:\\d{1,20}-)(\\d*)"))

# Extract chosen strand for primer design
if (grepl("primers",DNA_seq_header)==T) {
  # if 'primers' feature is indicated strand plus or minus is retrieved
  seq_strand <- stringr::str_extract(DNA_seq_header,"(?<=primers=)(plus|minus)")
  # if 'primers' feature doesn't indicate strand plus or minus : warning
  if (is.na(seq_strand)) { 
    seq_strand <- "plus"
    p("Warning: In fasta header 'primers' feature indicates neither plus or minus strand. Strand used for primers design is the plus strand by default.",
      style="color:#d37217 ; font-weight: bold ;")
  }
} else {
  # By default : plus strand
  seq_strand <- "plus"
  p("Warning: In fasta header no 'primers' feature is indicates. Strand used for primers design is the plus strand by default.", 
    style="color:#d37217 ; font-weight: bold ;")
}


# Load genome database for genomic plots
if (!(selected_genome %in% installed.packages()[,"Package"])) {
  BiocManager::install(selected_genome, ask = FALSE)
  suppressMessages(lapply(selected_genome, library, character.only = TRUE, quietly = TRUE))
}
genome <- getBSgenome(selected_genome)
```


```{r functions script,include=F}
# Functions script
source(file = "./scripts/ABSP_functions.R")
```


```{r output folders and file prefix}
# Create variable for common prefix of output files 
fileprefix <- paste(seq_name,seq_coll,seq_group,seq_rep,seq_clone,sep="_")
fileprefix <- gsub("__","_",fileprefix)
fileprefix <- gsub("*_$","",fileprefix)

ir_type_name <- paste0("individual_results_", ifelse(seq_clone!="","cloning","direct"))

# Create the results folder if not already created
if (! dir.exists(file.path(getwd(),"results"))) {suppressWarnings(dir.create(file.path(getwd(),"results")))}


# Path to main folder to seperate different experiment or user
indivDir <- list(
  main = file.path(getwd(),"results",folder_name))

# Path to sequence folder for the same sequence analyzed (same primers)
indivDir <- c(indivDir, list(
  name = file.path(indivDir$main,seq_name)))

# Paths to output folders
indivDir <- c(indivDir, list(
  indiv = file.path(indivDir$name,ir_type_name)))
indivDir <- c(indivDir, list(
  alg = file.path(indivDir$indiv,"alignments"),
  chrom = file.path(indivDir$indiv,"chromatograms"),
  data = file.path(indivDir$indiv,"data"),
  plot = file.path(indivDir$indiv,"plots"),
  seq = file.path(indivDir$indiv,"sequences"),
  tab = file.path(indivDir$indiv,"tables")
))
# Create directories
for (i in indivDir) {if (! dir.exists(i)){suppressWarnings(dir.create(i))}}



# Paths to output subfolders
indivDir$alg <- file.path(indivDir$alg,fileprefix)
indivDir$chrom <- file.path(indivDir$chrom,fileprefix)
indivDir$tab <- file.path(indivDir$tab,fileprefix)
# Create directories
for (i in indivDir) {if (! dir.exists(i)){suppressWarnings(dir.create(i))}}

```

```{r Diagram of output folders, inculde=F}
# create diagram of output folder with DiagrammeR
# settings
diagramset_indiv <- "
digraph graph2 { 

graph [layout = dot]

# node definitions with substituted label text
node [shape = folder, width = 1.5, fontname='Arial',fontsize=14, fillcolor='#374e64', style=filled, fontcolor='white',color=none, height=0.5]
a [label = 'ABSP']
b [label = 'results']
c [label = '@@1']
d [label = '@@2']
e [label = '@@3']
f [label = 'alignments']
g [label = 'chromatograms']
h [label = 'data']
i [label = 'plots']
j [label = 'sequences']
k [label = 'tables']
l [label = '.pair', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
m [label = '.fasta', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
n [label = '.pdf', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
o [label = '.csv', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
p [label = '.png', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
q [label = '.fasta', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
r [label = '.csv', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
s [label = '.png', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']
t [label = '.xlsx', shape= rectangle, style=rounded, width = 1,  fillcolor='white', fontcolor='#374e64',color='#374e64']

nodesep=0.2
ranksep=0.2
splines=polyline

a -> b [penwidth=0.5, arrowsize=0.5, arrowhead=none]
b -> c [penwidth=0.5, arrowsize=0.5, arrowhead=none]
c -> d [penwidth=0.5, arrowsize=0.5, arrowhead=none]
d -> e [penwidth=0.5, arrowsize=0.5, arrowhead=none]
e -> {f g h i j k} [penwidth=0.5, arrowhead=none]
f -> l [penwidth=0, arrowhead=none]
l -> m [penwidth=0, arrowhead=none]
g -> n [penwidth=0, arrowhead=none]
h -> o [penwidth=0, arrowhead=none]
i -> p [penwidth=0, arrowhead=none]
j -> q [penwidth=0, arrowhead=none]
k -> r [penwidth=0, arrowhead=none]
r -> s [penwidth=0, arrowhead=none]
s -> t [penwidth=0, arrowhead=none]

}
[1]:  paste0(folder_name)
[2]:  paste0(seq_name)
[3]:  paste0(ir_type_name)
"
# diagram
diagram_indiv <- DiagrammeR::grViz(diagramset_indiv)
```


```{r List the output files function, include=F}
list_output_indiv <- function(fileprefix) {
  files <- list(
    alg = list.files(path=indivDir$alg, full.names = T, include.dirs = F, recursive=T, pattern=fileprefix),
    chrom = list.files(path=indivDir$chrom, full.names = T, include.dirs = F, recursive=T, pattern=fileprefix),
    data = list.files(path=indivDir$data, full.names = T, include.dirs = F, recursive=T, pattern=fileprefix),
    plot = c(list.files(path=indivDir$plot, full.names = T, include.dirs = F, recursive=T, pattern=fileprefix),
             list.files(path=indivDir$plot, full.names = T, include.dirs = F, recursive=T, pattern="legend")),
    seq = list.files(path=indivDir$seq, full.names = T, include.dirs = F, recursive=T),
    tab = list.files(path=indivDir$tab, full.names = T, include.dirs = F, recursive=T, pattern=fileprefix))
  
  cat("  \n")
  if (length(files$alg)!=0) {
    cat("#### **alignments**","  \n")
    tags_alg <- tagList()
    for (i in files$alg) {tags_alg[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_alg))
    cat("  \n")}
  
  if (length(files$chrom)!=0) {
    cat("#### **chromatograms**","  \n")
    tags_chrom <- tagList()
    for (i in files$chrom) {tags_chrom[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_chrom))
    cat("  \n")}
  
  if (length(files$data)!=0) {
    cat("#### **data**","  \n")
    tags_data <- tagList()
    for (i in files$data) {tags_data[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_data))
    cat("  \n")}
  
  if (length(files$plot)!=0) {
    cat("#### **plots**","  \n")
    tags_plot <- tagList()
    for (i in files$plot) {tags_plot[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_plot))
    cat("  \n")}
  
  if (length(files$seq)!=0) {
    cat("#### **sequences**","  \n")
    tags_seq <- tagList()
    for (i in files$seq) {tags_seq[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_seq))
    cat("  \n")}
  
  if (length(files$tab)!=0) {
    cat("#### **tables**","  \n")
    tags_tab <- tagList()
    for (i in files$tab) {tags_tab[[i]] <- htmltools::p(htmltools::a(href = i,basename(i),target="_blank"))}
    cat(paste0(tags_tab))
    cat("  \n")}
}
```



```{r colors, include=F}
# Colors of sequence track for genomic plots
bases_colors <- c(A="#43CD80", T="#D7191C", G="#FDC661", C="#2C7BB6", N="#7F7F7F")
colors_tables <- c(correct="#00be7d",incorrect="#bf3232",warning="#d37217",correct_pale="#91e2c4",incorrect_pale="#e29191",warning_pale="#e2b891")
```





# {.tabset}

```{r display expermient parameters header, results='asis'}
# First column of table
parameters <- c(
  "Folder name :",
  "Sequence name :",
  "Collection :",
  "Group :",
  "Experiment type :",
  "Replicate number  :",
  "Clone number  :",
  "Date of sequencing #1 :",
  "Date of sequencing #2 :",
  "Date of analysis  :",
  "Prefix of output files  :"
  )
# Second column of table
values <- c(
  folder_name,
  seq_name,
  seq_coll,
  seq_group,
  "Type",
  seq_rep,
  seq_clone,
  seq_date_s1,
  seq_date_s2,
  format(Sys.time(), "%Y-%m-%d %H:%M"),
  fileprefix
)
# Specify the experiment type 
if(seq_rep!="") values[5] <- "Direct-BSP"
if(seq_clone!="") values[5] <- "Cloning-BSP"
# Create data frame
paramTable <- data.frame("Parameters" = parameters, "Input values" = values, check.names = F)
# Remove empty unecessary rows
paramTable <- paramTable[which(paramTable[,2]!=""),]
rownames(paramTable) <- NULL # reset row names, as row numbers are displayed if rows are missing
# print table as formattable
formattable(paramTable, align=rep("l", 2))
```


  

## Reference DNA {.tabset .tabset-pills}

### Reference DNA sequence {.tabset .tabset-pills}   

```{r choice of base DNA sequence from amplified strand, include=F}
# Get DNA sequences depending on strand chosen for primer design : 
  # base sequence is the sequence from the strand chosen for primer design as only this strand will be amplified by PCR following the bisulfite conversion
  # RC sequence is the reverse complement of base sequence, equivalent to the opposite strand, necessary to find CG coordinates on the reverse sequence (for CG position query on reverse sequencing results)
DNA_seq <- DNAString(DNA_seq)
if (seq_strand=="minus"){DNA_seq = reverseComplement(DNA_seq)}
seq <- list("seq" = DNA_seq, "seq_RC" = reverseComplement(DNA_seq))
```


| Reference DNA sequence information | Input values |
|:--------------------------|:--------------------------|
| Name                            | `r seq_name`                                    |
| Strand used for primer design   | `r stringr::str_to_sentence(seq_strand)` strand |
| Length                          | `r nchar(seq$seq)` pb                           |
| Coordinates on reference genome | `r paste0(seq_chr,":",seq_start,"-",seq_end) `  |




```{r check if incorrect genomic coordinates}
# Verify correct genomic coordinates
if ( (seq_end - seq_start) - length(DNA_seq) != -1 ) {genocoord <-T} else {genocoord <-F}
```
  
```{r output if incorrect genomic coordinates, results='asis', eval=genocoord}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if incorrect genomic coordinates, eval=genocoord}
# Diagram of directories
diagram_indiv
```

```{r stop knit if incorrect genomic coordinates, results='asis', eval=genocoord}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error: Length of reference DNA sequence does not match with the provided genomic coordinates. Please verify concordance between the reference DNA sequence and genomic coordinates.", style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1) 
knit_exit()
```





#### `r if(seq_strand=="plus"){"Genomic DNA Sequence from plus strand (template strand):"} else {"Genomic DNA Sequence from plus strand (not used):"} `
```{r display ref DNA sequence plus}
if(seq_strand=="plus") {
  cat(as.character(seq$seq))
  write.fasta(as.vector(seq$seq),
              names=paste(seq_name,"genomic DNA sequence from plus strand",paste0(seq_chr,":",seq_start,"-",seq_end),sep=" "),
              file.out=file.path(indivDir$seq,paste0(seq_name,"_plus.fasta")))} 
if(seq_strand=="minus") {
  cat(as.character(seq$seq_RC))
  write.fasta(as.vector(seq$seq_RC),
              names=paste(seq_name,"genomic DNA sequence from plus strand",paste0(seq_chr,":",seq_start,"-",seq_end),sep=" "),
              file.out=file.path(indivDir$seq,paste0(seq_name,"_plus.fasta")))}
```

#### `r if(seq_strand=="plus"){"Genomic DNA Sequence from minus strand (not used):"} else {"Genomic DNA Sequence from minus strand (template strand):"}`
```{r display ref DNA sequence minus}
if(seq_strand=="plus") {
  cat(as.character(seq$seq_RC))
  write.fasta(as.vector(seq$seq_RC),
              names=paste(seq_name,"genomic DNA sequence from minus strand",paste0(seq_chr,":",seq_start,"-",seq_end),sep=" "),
              file.out=file.path(indivDir$seq,paste0(seq_name,"_minus.fasta")))}
if(seq_strand=="minus") {
  cat(as.character(seq$seq))
  write.fasta(as.vector(seq$seq),
              names=paste(seq_name,"genomic DNA sequence from minus strand",paste0(seq_chr,":",seq_start,"-",seq_end),sep=" "),
              file.out=file.path(indivDir$seq,paste0(seq_name,"_minus.fasta")))}
```



### Localization of CG dinucleotides {.tabset .tabset-pills} 

```{r find CG localization, include=FALSE}
# Find CG coordinates from DNA ref coordinates
# To compute the coordinates, addition or removal of 1 is needed as the first nucleotide is corresponding to position 1 and not 0
CG_pos <- findCGcoordinates(sequence = seq$seq, 
                            sequence_reverse = seq$seq_RC, 
                            chr=seq_chr, coord_start = seq_start, coord_end = seq_end, 
                            seq_strand = seq_strand)
```


#### CG positions on plus strand

```{r display CG positions plus}
# Plus
if (seq_strand=="plus"){
  CG_pos_FT <- CG_pos$CG_pos
  colnames(CG_pos_FT) <- c("CG number","chr","start","end","C position","G position")
  formattable(CG_pos_FT, align=rep("c",ncol(CG_pos_FT)))
} 
if (seq_strand=="minus") {
  CG_pos_RC_FT <- CG_pos$CG_pos_RC
  colnames(CG_pos_RC_FT) <- c("CG number","chr","start","end","C position","G position")
  formattable(CG_pos_RC_FT, align=rep("c",ncol(CG_pos_RC_FT)))
}
```



#### CG positions on minus strand

```{r display CG positions minus}
# Minus
if (seq_strand=="plus") {
  CG_pos_RC_FT <- CG_pos$CG_pos_RC
  colnames(CG_pos_RC_FT) <- c("CG number","chr","start","end","C position","G position")
  formattable(CG_pos_RC_FT, align=rep("c",ncol(CG_pos_RC_FT)))
} 
if (seq_strand=="minus") {
  CG_pos_FT <- CG_pos$CG_pos
  colnames(CG_pos_FT) <- c("CG number","chr","start","end","C position","G position")
  formattable(CG_pos_FT, align=rep("c",ncol(CG_pos_FT)))
}
```



### Bisulfite converted sequences {.tabset .tabset-pills} 

<!--  The `r seq_strand` strand is used to perform bisulfite conversion as it is the strand used for primer design, only this strand is amplified during PCR.  
However, PCR regenerates the opposite strand of the bisulfite converted DNA sequence delimited by primers, which corresponds to its reverse complement but does not match with the bisulfite converted `r ifelse(seq_strand=="plus","minus","plus")` strand. -->
  

```{r bisulfite conversion}
# Bisulfite conversion is only performed on the chosen strand for primer design, as only this strand will be amplified by PCR following the bisulfite conversion
seq_BS <- bisulfiteConversion(seq$seq)
```

#### Converted sequence sense (template strand)  
Bisulfite converted sequence of the `r seq_strand` strand as indicated as the template strand.
```{r display DNA BS}
# Print and save the DNA sequence
cat(as.character(seq_BS$seq_BS))
fasta_name <- paste(seq_name, 
                    "bisulfite converted DNA sequence sense (from", seq_strand,"strand)", 
                    paste0(seq_chr, ":", seq_start, "-", seq_end), 
                    sep=" ")
write.fasta(as.vector(seq_BS$seq_BS), names=fasta_name, file.out=file.path(indivDir$seq,paste0(seq_name,"_BS_sense.fasta")))
rm(fasta_name)
```

#### Converted sequence antisense (reverse complement) 
Reverse complement of the bisulfite converted sequence from the `r seq_strand` strand.
```{r display DNA BS RC}
# Print and save the DNA sequence
cat(as.character(seq_BS$seq_BS_RC))
fasta_name <- paste(seq_name, 
                    "bisulfite converted DNA sequence antisense (from", seq_strand,"strand)", 
                    paste0(seq_chr, ":", seq_start, "-", seq_end), 
                    sep=" ")
write.fasta(as.vector(seq_BS$seq_BS_RC), names=fasta_name, file.out=file.path(indivDir$seq,paste0(seq_name,"_BS_antisense.fasta")))
rm(fasta_name)
```













## Sequencing trimming {.tabset .tabset-pills}

```{r Thresholds, include=F}

# Thresholds 

# Maximum base-calling error probability (value per position):
th_quality_error <- 0.001 
# Minimum Phred quality score, logarithmically linked to error probability (value per position):
th_quality_phred <- (-10*log(th_quality_error,10))

# Minimum ratio of primary peak, corresponding to the primary peak value over the total of peak value, to consider a position as non-mixed (value per position):
th_mixed_position <- 0.75
# Minimum percentage of non-mixed positions in the trimmed sequence to be considered non-mixed (value for the total trimmed sequence):
th_mixed_perc <- 75 # %

# Minimum length of trimmed sequences
th_min_trim <- 30 # bp

# Minimum length of the aligned sequences
th_min_alg <- 30 # bp
```

```{r read seq files, include=F}

shiny::setProgress(0.1) # For progression bar displayed

# SANGER SEQ FILES
if(useSeqs1==T) {
  sangerRead_s1 <- SangerRead(
    printLevel = "SangerRead",
    inputSource = "ABIF",
    readFeature = "Forward Read",
    readFileName = sangerseq_s1_path,
    geneticCode = GENETIC_CODE,
    TrimmingMethod = "M1",
    M1TrimmingCutoff = th_quality_error,
    M2CutoffQualityScore = NULL,
    M2SlidingWindowSize = NULL,
    baseNumPerRow = 2,
    heightPerRow = 200,
    signalRatioCutoff = (1-(th_mixed_perc/100)),
    showTrimmed = TRUE)
  
  if (sangerRead_s1@objectResults@creationResult==F) {useSeqs1 <- F}
}

if(useSeqs2==T) {
  sangerRead_s2 <- SangerRead(
    printLevel = "SangerRead",
    inputSource = "ABIF",
    readFeature = "Forward Read",
    readFileName = sangerseq_s2_path,
    geneticCode = GENETIC_CODE,
    TrimmingMethod = "M1",
    M1TrimmingCutoff = th_quality_error,
    M2CutoffQualityScore = NULL,
    M2SlidingWindowSize = NULL,
    baseNumPerRow = 2,
    heightPerRow = 200,
    signalRatioCutoff = (1-(th_mixed_perc/100)),
    showTrimmed = TRUE)
  
  if (sangerRead_s2@objectResults@creationResult==F) {useSeqs2 <- F}
}

shiny::setProgress(0.2) 
```

```{r create seq table s1, include=F, eval=useSeqs1}
# Create table
seqtable_s1 <- seqtable(sangerRead_s1)
```

```{r create seq table s2, include=F, eval=useSeqs2}
# Create table
seqtable_s2 <- seqtable(sangerRead_s2)
```


```{r mixed report s1, include=F, eval=useSeqs1}
# Find the trimmed sequence
ratios_s1 <- c()
for (i in 3:15) { # i consecutive non-mixed positions required at the start and end of the trimmed sequence
  MixedReport_s1 <- MixedReport(seqtable_s1,consecutive_threshold = i, th_mixed_position=th_mixed_position) 
  ratios_s1[i] <- MixedReport_s1$trimmedPercNonMixedPos
}
# select the minimum of consecutive positions requirement (to have the longer sequence) only above the ratio threshold (ratio of non-mixed over mixed position in the trimmed sequence)
th_mixed_consecutive_s1 <- ifelse(length(which(ratios_s1*100>=th_mixed_perc))>0 , min(which(ratios_s1*100>=th_mixed_perc)), NA)
if (!is.na(th_mixed_consecutive_s1)) {
  # Generate MixedReport for selected consecutive position threshold
  MixedReport_s1 <- MixedReport(seqtable_s1,consecutive_threshold = th_mixed_consecutive_s1, th_mixed_position=th_mixed_position)
  MixedReport_s1[["ConsecutiveNonMixedPosExtremities"]] <- th_mixed_consecutive_s1
} else {
  MixedReport_s1 <- MixedReport(seqtable_s1,consecutive_threshold = th_mixed_consecutive_s1, th_mixed_position=th_mixed_position)
  MixedReport_s1[["ConsecutiveNonMixedPosExtremities"]] <- th_mixed_consecutive_s1
}
```

```{r mixed report s2, include=F, eval=useSeqs2}
# Find the trimmed sequence
ratios_s2 <- c()
for (i in 3:15) { # i consecutive non-mixed positions required at the start and end of the trimmed sequence
  MixedReport_s2 <- MixedReport(seqtable_s2,consecutive_threshold = i, th_mixed_position=th_mixed_position) 
  ratios_s2[i] <- MixedReport_s2$trimmedPercNonMixedPos
}
# select the minimum of consecutive positions requirement (to have the longer sequence) only above the ratio threshold (ratio of non-mixed over mixed position in the trimmed sequence)
th_mixed_consecutive_s2 <- ifelse(length(which(ratios_s2*100>=th_mixed_perc))>0 , min(which(ratios_s2*100>=th_mixed_perc)), NA)
if (!is.na(th_mixed_consecutive_s2)) {
  # Generate MixedReport for selected consecutive position threshold
  MixedReport_s2 <- MixedReport(seqtable_s2,consecutive_threshold = th_mixed_consecutive_s2, th_mixed_position=th_mixed_position)
  MixedReport_s2[["ConsecutiveNonMixedPosExtremities"]] <- th_mixed_consecutive_s2
} else {
  MixedReport_s2 <- MixedReport(seqtable_s2,consecutive_threshold = th_mixed_consecutive_s2, th_mixed_position=th_mixed_position)
  MixedReport_s2[["ConsecutiveNonMixedPosExtremities"]] <- th_mixed_consecutive_s2
}
```



```{r trimmed sequence s1, results='asis', eval=useSeqs1}

# Trim sequencing result
Trimmed_s1 <- trimSeq(sangerRead_s1, MixedReport_s1)

if(Trimmed_s1$SeqLength < th_min_trim | is.na(Trimmed_s1$SeqLength)) {
  TrimOKs1 <- F
} else if (Trimmed_s1$MeanQualityScore < th_quality_phred | Trimmed_s1$PercNonMixedPos < th_mixed_perc ) { 
  TrimOKs1 <- F
} else {
  TrimOKs1 <- T
}
```

```{r trimmed sequence s2, results='asis', eval=useSeqs2}

# Trim sequencing result
Trimmed_s2 <- trimSeq(sangerRead_s2, MixedReport_s2)

if(Trimmed_s2$SeqLength <= th_min_trim | is.na(Trimmed_s2$SeqLength) ) {
  TrimOKs2 <- F
} else if (Trimmed_s2$MeanQualityScore < th_quality_phred | Trimmed_s2$PercNonMixedPos < th_mixed_perc ) { 
  TrimOKs2 <- F
} else {
  TrimOKs2 <- T
}
```



```{r final trimmed create table summary s1, eval=useSeqs1} 
# Create row dor sequencing 1
Trim_final_table_s1 <- data.frame(
  check.names = FALSE,
  "Sequencing" = "#1",
  "Length" = Trimmed_s1$SeqLength,
  "Start position" = Trimmed_s1$StartPos,
  "End position" = Trimmed_s1$FinishPo,
  "Mean of Phred quality scores" = Trimmed_s1$MeanQualityScore,
  "Mean of primary peak ratio" = Trimmed_s1$MeanPrimaryRatio,
  "Percentage of non-mixed positions" = Trimmed_s1$PercNonMixedPos,
  "Trimmed sequence quality" = TrimOKs1)
```

```{r final trimmed create table summary s2, eval=useSeqs2} 
# Create row dor sequencing 2
Trim_final_table_s2 <- data.frame(
  check.names = FALSE,
  "Sequencing" = "#2",
  "Length" = Trimmed_s2$SeqLength,
  "Start position" = Trimmed_s2$StartPos,
  "End position" = Trimmed_s2$FinishPo,
  "Mean of Phred quality scores" = Trimmed_s2$MeanQualityScore,
  "Mean of primary peak ratio" = Trimmed_s2$MeanPrimaryRatio,
  "Percentage of non-mixed positions" = Trimmed_s2$PercNonMixedPos,
  "Trimmed sequence quality" = TrimOKs2)
```

```{r final trimmed table summary}
# Merge tables 
if (useSeqs1==T & useSeqs2==T) {Trim_final_table <- rbind(Trim_final_table_s1, Trim_final_table_s2 )}
if (useSeqs1==T & useSeqs2==F) {Trim_final_table <- Trim_final_table_s1}
if (useSeqs1==F & useSeqs2==T) {Trim_final_table <- Trim_final_table_s2}
```




### Summary

| Thresholds                                  | Values                   |
|:-------------|:--------------------------|
| Minimum length of the trimmed sequence :        | **`r th_min_trim` bp**    |
| Minimum Phred quality score :  | **`r th_quality_phred`**    | 
| Minimum ratio of primary peak :  | **`r th_mixed_position`**    | 
| Minimum percentage of non-mixed positions :  | **`r th_mixed_perc` %**    | 


```{r final trimmed dispaly table summary, results='asis'}
# Export table data as csv and xlsx
write.table(Trim_final_table, file = file.path(indivDir$tab,paste0(fileprefix,"_trimmedseq_table.csv")), row.names = T, col.names = T)
write.xlsx(Trim_final_table, file = file.path(indivDir$tab,paste0(fileprefix,"_trimmedseq_table.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display Quality table
Trim_final_FT <- formattable(
  Trim_final_table, 
  align = rep("c",ncol(Trim_final_table)),
  list("Sequencing" = formatter("span", style = ~ formattable::style(font.weight = "bold")),
       "Length" = color_condi(">=", th_min_trim, colors_tables["correct_pale"], colors_tables["incorrect_pale"]),
       "Mean of Phred quality scores" =  color_condi(">=", th_quality_phred, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Mean of primary peak ratio" =  color_condi(">=", th_mixed_position, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Percentage of non-mixed positions" =  color_condi(">=", th_mixed_perc, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Trimmed sequence quality" = formatter("span", x ~ icontext(ifelse(x==T, "ok", "remove"), ifelse(x==T, "Correct", "Incorrect")), 
                                    style = x ~ formattable::style(color = ifelse(x==T, colors_tables["correct"], colors_tables["incorrect"])))))

Trim_final_FT
```

```{r Trim summary export table, include=F}
export_formattable(Trim_final_FT,file=file.path(indivDir$tab,paste0(fileprefix,"_trimmedseq.png")))
```


### Sequencing result #1 {.tabset .tabset-pills}

#### Raw sequence {.tabset .tabset-pills}

##### Sequence

```{r raw seq s1, results='asis', eval=useSeqs1}
p("Sequence of primary peaks:", style="font-weight: bold ;")
p("Length: ",nchar(primarySeq(sangerRead_s1,string = TRUE))," bp")
```

```{r display primary peaks seq s1, eval=useSeqs1}
cat(primarySeq(sangerRead_s1,string = TRUE))
```

##### Chromatogram
```{r chromatogram s1, include=F, eval=useSeqs1}
# Save chromatogram
chroma_s1_path <- file.path(indivDir$chrom,paste0(fileprefix,"_chromatogram_1.pdf"))
chromatogram(sangerRead_s1, width = 20, height = 2, showcalls = "both", filename = chroma_s1_path)
```

```{r display chromatogram s1, out.width = "100%", fig.align = "center", eval=useSeqs1}
htmltools::p("Open ",htmltools::a(href = file.path(chroma_s1_path),"chromatogram file",target="_blank")," in a new tab")
htmltools::tags$iframe(src=file.path(chroma_s1_path), style="width:100%; height:1000px;", frameborder="0")
```

##### Table
```{r seq table s1, include=F, eval=useSeqs1}
# Export table data as csv and xlsx
write.table(seqtable_s1, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_1.csv")), row.names = T, col.names = T)
write.xlsx(seqtable_s1, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_1.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display table
seqtable_s1_FT <- seqtable_s1
colnames(seqtable_s1_FT) <- c("Position","A peak","T peak","G peak","C peak","Total","A ratio","T ratio","G ratio","C ratio","Primary base","Quality score","Primary ratio")
seqtable_s1_FT <- formattable(
  seqtable_s1_FT, 
  align = rep("c",ncol(seqtable_s1_FT)),
  list("Position" = formatter("span", style = ~ formattable::style(font.weight = "bold")), 
       "A peak"= color_format("white",bases_colors["A"]),
       "T peak"= color_format("white",bases_colors["T"]),
       "G peak"= color_format("white",bases_colors["G"]),
       "C peak"= color_format("white",bases_colors["C"]),
       "Total"= color_condi(">=", 100, colorT = "lightgrey", colorF = colors_tables["incorrect_pale"],block="block"),
       "A ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["A"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "T ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["T"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "G ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["G"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "C ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["C"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "Primary base"= color_base("white", bases_colors=bases_colors),
       "Quality score" = color_condi(">=", th_quality_phred,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block"),
       "Primary ratio" = color_condi(">=", th_mixed_position,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block")
  ))
```

```{r display seq table s1, eval=useSeqs1}
seqtable_s1_FT
```


#### Quality report


**Trimming thresholds :**  

* Trimming Method : **`r ifelse(useSeqs1==T,sangerRead_s1@QualityReport@TrimmingMethod,"")`**  
* Minimum base error probability : **`r ifelse(useSeqs1==T,th_quality_error,"")` %**  
* Minimum Phred quality score : **`r ifelse(useSeqs1==T,th_quality_phred,"")`**  


| Raw sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Raw sequence length : | **`r ifelse(useSeqs1==T,sangerRead_s1@QualityReport@rawSeqLength,"")` bp** |
| Mean of raw sequence Phred quality scores : | **`r ifelse(useSeqs1==T,round(sangerRead_s1@QualityReport@rawMeanQualityScore,2),"")`** |


| Quality report - Trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Trimmed sequence length : | **`r ifelse(useSeqs1==T,sangerRead_s1@QualityReport@trimmedSeqLength,"")` bp** |
| Trimmed sequence start position : | **`r ifelse(useSeqs1==T,sangerRead_s1@QualityReport@trimmedStartPos,"")`** |
| Trimmed sequence end position : | **`r ifelse(useSeqs1==T,sangerRead_s1@QualityReport@trimmedFinishPos,"")`** |
| Mean of trimmed sequence Phred quality scores : | **`r ifelse(useSeqs1==T,round(sangerRead_s1@QualityReport@trimmedMeanQualityScore,2),"")`** |
 

#### Mixed base peak report

 
**Trimming thresholds :**  

* Minimum primary peak ratio for a position to be considered non-mixed : **`r ifelse(useSeqs1==T,th_mixed_position,"")`**  
* Minimum of non-mixed positions percentage in the trimmed sequence to be considered non-mixed : **`r ifelse(useSeqs1==T,th_mixed_perc,"")` %**  
* Minimum number of consecutive non-mixed positions at extremities to get a non-mixed positions percentage above threshold for the trimmed sequence (computed) : **`r ifelse(useSeqs1==T,MixedReport_s1$ConsecutiveNonMixedPosExtremities,"")` bp**  


| Raw sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Raw sequence length : | **`r ifelse(useSeqs1==T,MixedReport_s1$rawSeqLength,"")` bp** |
| Mean of raw sequence primary peak ratios : | **`r ifelse(useSeqs1==T, MixedReport_s1$rawMeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions on raw sequence : | **`r ifelse(useSeqs1==T,MixedReport_s1$rawPercNonMixedPos,"")` %** |


| Mixed peak report - Trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Trimmed sequence length : | **`r ifelse(useSeqs1==T,MixedReport_s1$trimmedSeqLength,"")` bp** |
| Trimmed sequence start position : | **`r ifelse(useSeqs1==T,MixedReport_s1$trimmedStartPos,"")`** |
| Trimmed sequence end position : | **`r ifelse(useSeqs1==T,MixedReport_s1$trimmedFinishPos,"")`** |
| Mean of trimmed sequence primary peak ratios : | **`r ifelse(useSeqs1==T,MixedReport_s1$trimmedMeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions on trimmed sequence : | **`r ifelse(useSeqs1==T,MixedReport_s1$trimmedPercNonMixedPos,"")` %** |


#### Trimming plot 

```{r trim plot s1, include=F, eval=useSeqs1}
Trim_plot_s1 <- suppressWarnings(trimming_plot(sangerRead = sangerRead_s1, 
                                               seqtable = seqtable_s1, 
                                               MixedReport = MixedReport_s1, 
                                               Trimmed_seq = Trimmed_s1,
                                               th_quality_phred = th_quality_phred, 
                                               th_mixed_position = th_mixed_position,
                                               downloadname = paste0(fileprefix,"_trimming_plot_s1.png"),
                                               colors_tables = colors_tables))
```

```{r display trim plot s1, out.width="100%", eval=useSeqs1}
suppressWarnings(Trim_plot_s1)
```




#### Final trimmed sequence {.tabset .tabset-pills}

##### Sequence

```{r trimmed ok s1, results='asis', eval=useSeqs1 & !TrimOKs1}
p("Warning: Trimmed sequence from the sequencing result #1 was found with insufficient length or insufficient quality and will not be used.",
  style="color:#d37217 ; font-weight: bold ;")
```


| Final trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Length : | **`r ifelse(useSeqs1==T,Trimmed_s1$SeqLength,"")` bp** |
| Start position : | **`r ifelse(useSeqs1==T,Trimmed_s1$StartPos,"")`** |
| End position : | **`r ifelse(useSeqs1==T,Trimmed_s1$FinishPos,"")`** |
| Mean of Phred quality scores : | **`r ifelse(useSeqs1==T,Trimmed_s1$MeanQualityScore,"")`** |
| Mean of primary peak ratio : | **`r ifelse(useSeqs1==T,Trimmed_s1$MeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions : | **`r ifelse(useSeqs1==T,Trimmed_s1$PercNonMixedPos,"")` %** |
  
  
```{r trimmed seq s1, results='asis', eval=useSeqs1 & TrimOKs1}
p("Sequence of primary peaks:", style="font-weight: bold ;")
```

```{r display trimmed seq s1, eval=useSeqs1 & TrimOKs1}
cat(as.character(Trimmed_s1$PrimarySeq))
```


##### Chromatogram

```{r chromatogram trimmed s1, include=F, eval=useSeqs1 & TrimOKs1}
# Save chromatogram
chroma_s1_path <- file.path(indivDir$chrom,paste0(fileprefix,"_chromatogram_trimmed_1.pdf"))
chromatogram(sangerRead_s1, trim5 = Trimmed_s1$StartPos, trim3 = (sangerRead_s1@QualityReport@rawSeqLength - Trimmed_s1$FinishPos), width = 20, height = 2, showcalls = "both", filename = chroma_s1_path)
```

```{r display trimmed chromatogram s1, out.width = "100%", fig.align = "center", eval=useSeqs1 & TrimOKs1}

htmltools::p("Open ",htmltools::a(href = file.path(chroma_s1_path),"chromatogram file",target="_blank")," in a new tab")
htmltools::tags$iframe(src=file.path(chroma_s1_path), style="width:100%; height:1000px;", frameborder="0") 
```



##### Table

```{r trimmed seq table s1,include=F, eval=useSeqs1 & TrimOKs1}

# Create table
seqtable_trim_s1 <- seqtable_s1[Trimmed_s1$StartPos:Trimmed_s1$FinishPos,]

# Export table data as csv and xlsx
write.table(seqtable_trim_s1, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_trimmed_1.csv")), row.names = T, col.names = T)
write.xlsx(seqtable_trim_s1, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_trimmed_1.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display table
seqtable_trim_s1_FT <- seqtable_trim_s1
colnames(seqtable_trim_s1_FT) <- c("Position","A peak","T peak","G peak","C peak","Total","A ratio","T ratio","G ratio","C ratio","Primary base","Quality score","Primary ratio")
seqtable_trim_s1_FT <- formattable(
  seqtable_trim_s1_FT, 
  align = rep("c",ncol(seqtable_trim_s1_FT)),
  list("Position" = formatter("span", style = ~ formattable::style(font.weight = "bold")), 
       "A peak"= color_format("white",bases_colors["A"]),
       "T peak"= color_format("white",bases_colors["T"]),
       "G peak"= color_format("white",bases_colors["G"]),
       "C peak"= color_format("white",bases_colors["C"]),
       "Total"= color_condi(">=", 100, colorT = "lightgrey", colorF = colors_tables["incorrect_pale"],block="block"),
       "A ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["A"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "T ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["T"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "G ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["G"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "C ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["C"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "Primary base"= color_base("white", bases_colors=bases_colors),
       "Quality score" = color_condi(">=", th_quality_phred,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block"),
       "Primary ratio" = color_condi(">=", th_mixed_position,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block")
  ))
```

```{r display trimmed seq table s1, eval=useSeqs1 & TrimOKs1}
seqtable_trim_s1_FT

shiny::setProgress(0.3) 
```










### Sequencing result #2 {.tabset .tabset-pills}

#### Raw sequence {.tabset .tabset-pills}

##### Sequence

```{r raw seq s2, results='asis', eval=useSeqs2}
p("Sequence of primary peaks:", style="font-weight: bold ;")
p("Length: ",nchar(primarySeq(sangerRead_s2,string = TRUE))," bp")
```

```{r display primary peaks seq s2, eval=useSeqs2}
cat(primarySeq(sangerRead_s2,string = TRUE))
```

##### Chromatogram
```{r chromatogram s2, include=F, eval=useSeqs2}
# Save chromatogram
chroma_s2_path <- file.path(indivDir$chrom,paste0(fileprefix,"_chromatogram_2.pdf"))
chromatogram(sangerRead_s2, width = 20, height = 2, showcalls = "both", filename = chroma_s2_path)
```

```{r display chromatogram s2, out.width = "100%", fig.align = "center", eval=useSeqs2}
htmltools::p("Open ",htmltools::a(href = file.path(chroma_s2_path),"chromatogram file",target="_blank")," in a new tab")
htmltools::tags$iframe(src=file.path(chroma_s2_path), style="width:100%; height:1000px;", frameborder="0")
```

##### Table
```{r seq table s2, include=F, eval=useSeqs2}
# Export table data as csv and xlsx
write.table(seqtable_s2, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_2.csv")), row.names = T, col.names = T)
write.xlsx(seqtable_s2, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_2.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display table
seqtable_s2_FT <- seqtable_s2
colnames(seqtable_s2_FT) <- c("Position","A peak","T peak","G peak","C peak","Total","A ratio","T ratio","G ratio","C ratio","Primary base","Quality score","Primary ratio")
seqtable_s2_FT <- formattable(
  seqtable_s2_FT, 
  align = rep("c",ncol(seqtable_s2_FT)),
  list("Position" = formatter("span", style = ~ formattable::style(font.weight = "bold")), 
       "A peak"= color_format("white",bases_colors["A"]),
       "T peak"= color_format("white",bases_colors["T"]),
       "G peak"= color_format("white",bases_colors["G"]),
       "C peak"= color_format("white",bases_colors["C"]),
       "Total"= color_condi(">=", 100, colorT = "lightgrey", colorF = colors_tables["incorrect_pale"],block="block"),
       "A ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["A"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "T ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["T"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "G ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["G"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "C ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["C"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "Primary base"= color_base("white", bases_colors=bases_colors),
       "Quality score" = color_condi(">=", th_quality_phred,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block"),
       "Primary ratio" = color_condi(">=", th_mixed_position,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block")
  ))
```

```{r display seq table s2, eval=useSeqs2}
seqtable_s2_FT
```


#### Quality report


**Trimming thresholds :**  

* Trimming Method : **`r ifelse(useSeqs2==T,sangerRead_s2@QualityReport@TrimmingMethod,"")`**  
* Minimum base error probability : **`r ifelse(useSeqs2==T,th_quality_error,"")` %**  
* Minimum Phred quality score : **`r ifelse(useSeqs2==T,th_quality_phred,"")`**  


| Raw sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Raw sequence length : | **`r ifelse(useSeqs2==T,sangerRead_s2@QualityReport@rawSeqLength,"")` bp** |
| Mean of raw sequence Phred quality scores : | **`r ifelse(useSeqs2==T,round(sangerRead_s2@QualityReport@rawMeanQualityScore,2),"")`** |


| Quality report - Trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Trimmed sequence length : | **`r ifelse(useSeqs2==T,sangerRead_s2@QualityReport@trimmedSeqLength,"")` bp** |
| Trimmed sequence start position : | **`r ifelse(useSeqs2==T,sangerRead_s2@QualityReport@trimmedStartPos,"")`** |
| Trimmed sequence end position : | **`r ifelse(useSeqs2==T,sangerRead_s2@QualityReport@trimmedFinishPos,"")`** |
| Mean of trimmed sequence Phred quality scores : | **`r ifelse(useSeqs2==T,round(sangerRead_s2@QualityReport@trimmedMeanQualityScore,2),"")`** |
 


#### Mixed base peak report


**Trimming thresholds :**  

* Minimum primary peak ratio for a position to be considered non-mixed : **`r ifelse(useSeqs2==T,th_mixed_position,"")`**  
* Minimum of non-mixed positions percentage in the trimmed sequence to be considered non-mixed : **`r ifelse(useSeqs2==T,th_mixed_perc,"")` %**  
* Minimum number of consecutive non-mixed positions at extremities to get a non-mixed positions percentage above threshold for the trimmed sequence (computed) : **`r ifelse(useSeqs2==T,MixedReport_s2$ConsecutiveNonMixedPosExtremities,"")` bp**  


| Raw sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Raw sequence length : | **`r ifelse(useSeqs2==T,MixedReport_s2$rawSeqLength,"")` bp** |
| Mean of raw sequence primary peak ratios : | **`r ifelse(useSeqs2==T, MixedReport_s2$rawMeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions on raw sequence : | **`r ifelse(useSeqs2==T,MixedReport_s2$rawPercNonMixedPos,"")` %** |


| Mixed peak report - Trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Trimmed sequence length : | **`r ifelse(useSeqs2==T,MixedReport_s2$trimmedSeqLength,"")` bp** |
| Trimmed sequence start position : | **`r ifelse(useSeqs2==T,MixedReport_s2$trimmedStartPos,"")`** |
| Trimmed sequence end position : | **`r ifelse(useSeqs2==T,MixedReport_s2$trimmedFinishPos,"")`** |
| Mean of trimmed sequence primary peak ratios : | **`r ifelse(useSeqs2==T,MixedReport_s2$trimmedMeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions on trimmed sequence : | **`r ifelse(useSeqs2==T,MixedReport_s2$trimmedPercNonMixedPos,"")` %** |



#### Trimming plot 

```{r trim plot s2, include=F, eval=useSeqs2}
Trim_plot_s2 <- suppressWarnings(trimming_plot(sangerRead = sangerRead_s2, 
                                               seqtable = seqtable_s2, 
                                               MixedReport = MixedReport_s2, 
                                               Trimmed_seq = Trimmed_s2,
                                               th_quality_phred = th_quality_phred, 
                                               th_mixed_position = th_mixed_position,
                                               downloadname = paste0(fileprefix,"_trimming_plot_s2.png"),
                                               colors_tables = colors_tables))
```

```{r display trim plot s2, out.width="100%", eval=useSeqs2}
suppressWarnings(Trim_plot_s2)
```




#### Final trimmed sequence {.tabset .tabset-pills}

##### Sequence

```{r trimmed ok s2, results='asis', eval=useSeqs2 & !TrimOKs2}
p("Warning: Trimmed sequence from the sequencing result #2 was found with insufficient length or insufficient quality and will not be used.",
  style="color:#d37217 ; font-weight: bold ;")
```



| Final trimmed sequence information | Values |
|:----------------------------------------------------|:-----------------------|
| Length : | **`r ifelse(useSeqs2==T,Trimmed_s2$SeqLength,"")` bp** |
| Start position : | **`r ifelse(useSeqs2==T,Trimmed_s2$StartPos,"")`** |
| End position : | **`r ifelse(useSeqs2==T,Trimmed_s2$FinishPos,"")`** |
| Mean of Phred quality scores : | **`r ifelse(useSeqs2==T,Trimmed_s2$MeanQualityScore,"")`** |
| Mean of primary peak ratio : | **`r ifelse(useSeqs2==T,Trimmed_s2$MeanPrimaryRatio,"")`** |
| Percentage of non-mixed positions : | **`r ifelse(useSeqs2==T,Trimmed_s2$PercNonMixedPos,"")` %** |


```{r trimmed seq s2, results='asis', eval=useSeqs2 & TrimOKs2}
p("Sequence of primary peaks:", style="font-weight: bold ;")
```

```{r display trimmed seq s2, eval=useSeqs2 & TrimOKs2}
cat(as.character(Trimmed_s2$PrimarySeq))
```



##### Chromatogram

```{r chromatogram trimmed s2, include=F, eval=useSeqs2 & TrimOKs2}
# Save chromatogram
chroma_s2_path <- file.path(indivDir$chrom,paste0(fileprefix,"_chromatogram_trimmed_2.pdf"))
chromatogram(sangerRead_s2, trim5 = Trimmed_s2$StartPos, trim3 = (sangerRead_s2@QualityReport@rawSeqLength - Trimmed_s2$FinishPos), width = 20, height = 2, showcalls = "both", filename = chroma_s2_path) 
```

```{r display trimmed chromatogram s2, out.width = "100%", fig.align = "center", eval=useSeqs2 & TrimOKs2}
htmltools::p("Open ",htmltools::a(href = file.path(chroma_s2_path),"chromatogram file",target="_blank")," in a new tab")
htmltools::tags$iframe(src=file.path(chroma_s2_path), style="width:100%; height:1000px;", frameborder="0") 
```



##### Table

```{r trimmed seq table s2,include=F, eval=useSeqs2 & TrimOKs2}

# Create table
seqtable_trim_s2 <- seqtable_s2[Trimmed_s2$StartPos:Trimmed_s2$FinishPos,]

# Export table data as csv and xlsx
write.table(seqtable_trim_s2, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_trimmed_2.csv")), row.names = T, col.names = T)
write.xlsx(seqtable_trim_s2, file = file.path(indivDir$tab,paste0(fileprefix,"_seqtable_trimmed_2.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display table
seqtable_trim_s2_FT <- seqtable_trim_s2
colnames(seqtable_trim_s2_FT) <- c("Position","A peak","T peak","G peak","C peak","Total","A ratio","T ratio","G ratio","C ratio","Primary base","Quality score","Primary ratio")
seqtable_trim_s2_FT <- formattable(
  seqtable_trim_s2_FT, 
  align = rep("c",ncol(seqtable_trim_s2_FT)),
  list("Position" = formatter("span", style = ~ formattable::style(font.weight = "bold")), 
       "A peak"= color_format("white",bases_colors["A"]),
       "T peak"= color_format("white",bases_colors["T"]),
       "G peak"= color_format("white",bases_colors["G"]),
       "C peak"= color_format("white",bases_colors["C"]),
       "Total"= color_condi(">=", 100, colorT = "lightgrey", colorF = colors_tables["incorrect_pale"],block="block"),
       "A ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["A"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "T ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["T"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "G ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["G"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "C ratio"= color_condi(">=", th_mixed_position,colorT = bases_colors["C"],colorF = "white",textcolorT = "white",textcolorF = ""),
       "Primary base"= color_base("white", bases_colors=bases_colors),
       "Quality score" = color_condi(">=", th_quality_phred,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block"),
       "Primary ratio" = color_condi(">=", th_mixed_position,colorT = colors_tables["correct_pale"],colorF = colors_tables["incorrect_pale"],block="block")
  )) 
```

```{r display trimmed seq table s2, eval=useSeqs2 & TrimOKs2}
seqtable_trim_s2_FT

shiny::setProgress(0.4) 
```





```{r check if no trim possible, include=F}
# Verify if no CG covered by both sequencing
if (TrimOKs1==F & TrimOKs2==F) {NoTrimmedSeq <- T} else {NoTrimmedSeq <- F}
```
  
```{r output if no trimmed sequence, results='asis', eval=NoTrimmedSeq}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if no trimmed sequence, eval=NoTrimmedSeq}
# Diagram of directories
diagram_indiv
```

```{r stop knit if no trimmed sequence, results='asis', eval=NoTrimmedSeq}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error:  Analysis has been stopped as none of the sequencings are of sufficient quality to be used.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1)
knit_exit()
```






## Alignments {.tabset .tabset-pills}


```{r alignments s1, eval=useSeqs1 & TrimOKs1}
# Alignments #1

pa_s1_S <- pairwiseAlignment(Trimmed_s1$PrimarySeq, seq_BS$seq_BS, type = "local")
pa_s1_RC <- pairwiseAlignment(Trimmed_s1$PrimarySeq, seq_BS$seq_BS_RC, type = "local")

if (pa_s1_S@subject@range@width < th_min_alg & 
    pa_s1_RC@subject@range@width < th_min_alg) {NoAlignmts1 <- T} else {NoAlignmts1 <- F}

```

```{r alignments s2, eval=useSeqs2 & TrimOKs2}
# Alignments #2
pa_s2_S <- pairwiseAlignment(Trimmed_s2$PrimarySeq, seq_BS$seq_BS, type = "local")
pa_s2_RC <- pairwiseAlignment(Trimmed_s2$PrimarySeq, seq_BS$seq_BS_RC, type = "local")

if (pa_s2_S@subject@range@width < th_min_alg & 
    pa_s2_RC@subject@range@width < th_min_alg) {NoAlignmts2 <- T} else {NoAlignmts2 <- F}
```



```{r check if alignments, include=F, eval=useSeqs1 & useSeqs2}
# Verify if alignments of correct trimmed sequence are long enough
if(TrimOKs1==T & TrimOKs2==T) {
  if(NoAlignmts1==T & NoAlignmts2==T) {NoAlignmt <- T} else {NoAlignmt <- F}
}
if(TrimOKs1==T & TrimOKs2==F) {
  if(NoAlignmts1==T) {NoAlignmt <- T} else {NoAlignmt <- F}
}
if(TrimOKs1==F & TrimOKs2==T) {
  if(NoAlignmts2==T) {NoAlignmt <- T} else {NoAlignmt <- F}
}
```

```{r check if alignments bis, include = F}
if(useSeqs2==F & TrimOKs1==T) { if(NoAlignmts1==T) {NoAlignmt <- T} else {NoAlignmt <- F} }
if(useSeqs1==F & TrimOKs2==T) { if(NoAlignmts2==T) {NoAlignmt <- T} else {NoAlignmt <- F} }
```

```{r output if alignments too shorts, results='asis', eval=NoAlignmt}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if alignments too shorts, eval=NoAlignmt}
# Diagram of directories
diagram_indiv
```

```{r stop knit if alignments too shorts, results='asis', eval=NoAlignmt}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error:  Analysis has been stopped as none of the possible alignments are of sufficient length (< ",th_min_alg," bp) to be used.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1)
knit_exit()
```




### Sequencing result #1 {.tabset .tabset-pills}

Sequencing result #1 is aligned on both sense and antisense bisulfite converted DNA sequences to determine the sequencing direction as forward or reverse.

```{r text warning no trim s1, eval=useSeqs1 & !TrimOKs1}
p("Warning: Trimmed sequence from the sequencing result #1 was found with insufficient length or insufficient quality and will not be used.",
  style="color:#d37217 ; font-weight: bold ;")
```



#### Alignment as forward

Trimmed sequence of primary peaks from sequencing result **#1** is aligned with bisulfite converted reference DNA **sense** sequence
```{r s1 forward, eval=useSeqs1 & TrimOKs1}
writePairwiseAlignments(pa_s1_S,block.width=20)
```



#### Alignment as reverse

Trimmed sequence of primary peaks from sequencing result **#1** is aligned with bisulfite converted reference DNA **antisense** sequence
```{r s1 reverse, eval=useSeqs1 & TrimOKs1}
writePairwiseAlignments(pa_s1_RC,block.width=20)
```



#### Direction determination
```{r direction determination s1, include=F, eval=useSeqs1}
#
# RESULT 1 : USE ALIGNMENTS LENGHT TO DETERMINE CORRECT DIRECTION
if(TrimOKs1==T) {
  if(pa_s1_S@subject@range@width > th_min_alg | pa_s1_RC@subject@range@width > th_min_alg) {
    
    # RESULT 1 ALIGNED WITH SENSE SEQUENCE : FORWARD
    if (pa_s1_S@subject@range@width > pa_s1_RC@subject@range@width) {
      # Result 1 is in forward direction
      useSeqF <- useSeqs1
      TrimOKF <- TrimOKs1
      seqnumber_F <- 1
      seqdirection_s1 <- "Forward"
    }
    
    # RESULT 1 ALIGNED WITH ANTISENSE SEQUENCE : REVERSE
    if (pa_s1_S@subject@range@width < pa_s1_RC@subject@range@width) {
      # Result 1 is in reverse direction
      useSeqR <- useSeqs1
      TrimOKR <- TrimOKs1
      seqnumber_R <- 1
      seqdirection_s1 <- "Reverse"
    }
    DIRincorrect <- F # Direction determination is sucessfull
    Dirdeterms2 <- F # not use s2 for direction determination
  } else {
    Dirdeterms2 <- T # use s2 for direction determination when aligned length are too small
  }
  
  
  if (pa_s1_S@subject@range@width == pa_s1_RC@subject@range@width) {
    Dirdeterms2 <- T # use s2 for direction determination when both length are equal
  } 
} else {
  Dirdeterms2 <- T # use s2 for direction determination when trimming is not correct
} 

#
# RESULT 1 NO DIRECTION DETERMINATION POSSIBLE : USE RESULT 2
if(Dirdeterms2==T) {
  if (useSeqs2==T & TrimOKs2==T) {
    if (pa_s2_S@subject@range@width > pa_s2_RC@subject@range@width) {
      # Result 2 is in forward direction
      # Result 1 is in reverse direction
      useSeqR <- useSeqs1
      TrimOKR <- TrimOKs1
      seqnumber_R <- 2
      seqdirection_s1 <- "Reverse"
    }
    if (pa_s2_S@subject@range@width < pa_s2_RC@subject@range@width) {
      # Result 2 is in reverse direction
      # Result 1 is in forward direction
      useSeqF <- useSeqs1
      TrimOKF <- TrimOKs1
      seqnumber_F <- 2
      seqdirection_s1 <- "Forward"
    }
    DIRincorrect <- F # Direction determination is sucessfull
    # RESULT 2 UNDETERMINED ALIGNMENT : WARNING
  } else if (useSeqs2==F | TrimOKs2==F) {
    p("Warning: Sequencing result #1 direction undetermined, aligned sequences length are equal for both directions.",
      style="color:#d37217 ; font-weight: bold ;")
    DIRincorrect <- T # Direction determination is not sucessfull
    seqdirection_s1 <- "undetermined"
  }
}
```

```{r text direction determination s1, results='asis', eval=useSeqs1}
if(seqdirection_s1=="Forward") {
  if(TrimOKF==T) {
    cat("Sequencing result **#1** corresponds to **forward** direction")
    pa_F <- pa_s1_S
    seqtable_trim_F <- seqtable_trim_s1
  } else if (TrimOKF==F) {
    p("Warning: Insufficient quality of trimmed sequence for direction determination.",
      style="color:#d37217 ; font-weight: bold ;")
  }
}

if(seqdirection_s1=="Reverse") {
  if(TrimOKR==T) {
    cat("Sequencing result **#1** corresponds to **reverse** direction")
    pa_R <- pa_s1_RC
    seqtable_trim_R <- seqtable_trim_s1
  } else if (TrimOKR==F) {
    p("Warning: Insufficient quality of trimmed sequence for direction determination.",
      style="color:#d37217 ; font-weight: bold ;")
  }
}
```






### Sequencing result #2 {.tabset .tabset-pills}

Sequencing result #2 is aligned on both sense and antisense bisulfite converted DNA sequences to determine the sequencing direction as forward or reverse.

```{r text warning no trim s2, eval=useSeqs2 & !TrimOKs2}
p("Warning: Trimmed sequence from the sequencing result #2 was found with insufficient length or insufficient quality and will not be used.",
  style="color:#d37217 ; font-weight: bold ;")
```


#### Alignment as forward

Trimmed sequence of primary peaks from sequencing result **#2** is aligned with bisulfite converted reference DNA **sense** sequence
```{r s2 forward, eval=useSeqs2 & TrimOKs2}
writePairwiseAlignments(pa_s2_S,block.width=20)
```



#### Alignment as reverse

Trimmed sequence of primary peaks from sequencing result **#2** is aligned with bisulfite converted reference DNA **antisense** sequence
```{r s2 reverse, eval=useSeqs2 & TrimOKs2}
writePairwiseAlignments(pa_s2_RC,block.width=20)
```



#### Direction determination
```{r direction determination s2, include=F, eval=useSeqs2}
#
# RESULT 2 : USE ALIGNMENTS LENGHT TO DETERMINE CORRECT DIRECTION
if(TrimOKs2==T) {
  if(pa_s2_S@subject@range@width > th_min_alg | pa_s2_RC@subject@range@width > th_min_alg) {
    
    # RESULT 2 ALIGNED WITH SENSE SEQUENCE : FORWARD
    if (pa_s2_S@subject@range@width > pa_s2_RC@subject@range@width) {
      # Result 2 is in forward direction
      useSeqF <- useSeqs2
      TrimOKF <- TrimOKs2
      seqnumber_F <- 2
      seqdirection_s2 <- "Forward"
    }
    
    # RESULT 2 ALIGNED WITH ANTISENSE SEQUENCE : REVERSE
    if (pa_s2_S@subject@range@width < pa_s2_RC@subject@range@width) {
      # Result 2 is in reverse direction
      useSeqR <- useSeqs2
      TrimOKR <- TrimOKs2
      seqnumber_R <- 2
      seqdirection_s2 <- "Reverse"
    }
    DIRincorrect <- F # Direction determination is sucessfull
    Dirdeterms1 <- F # not use s1 for direction determination
  } else {
    Dirdeterms1 <- T # use s1 for direction determination when aligned length are too small
  }
  
  if (pa_s2_S@subject@range@width == pa_s2_RC@subject@range@width) {
    Dirdeterms1 <- T # use s1 for direction determination when both length are equal
  } 
} else {
  Dirdeterms1 <- T # use s1 for direction determination when trimming is not correct
} 

#
# RESULT 2 NO DIRECTION DETERMINATION POSSIBLE : USE RESULT 1
if(Dirdeterms1==T) {
  if (useSeqs1==T & TrimOKs1==T) {
    if (pa_s1_S@subject@range@width > pa_s1_RC@subject@range@width) {
      # Result 1 is in forward direction
      # Result 2 is in reverse direction
      useSeqR <- useSeqs2
      TrimOKR <- TrimOKs2
      seqnumber_R <- 2
      seqdirection_s2 <- "Reverse"
    }
    if (pa_s1_S@subject@range@width < pa_s1_RC@subject@range@width) {
      # Result 1 is in reverse direction
      # Result 2 is in forward direction
      useSeqF <- useSeqs2
      TrimOKF <- TrimOKs2
      seqnumber_F <- 2
      seqdirection_s2 <- "Forward"
    }
    DIRincorrect <- F # Direction determination is sucessfull
    # RESULT 2 UNDETERMINED ALIGNMENT : WARNING
  } else if (useSeqs1==F | TrimOKs1==F) {
    p("Warning: Sequencing result #2 direction undetermined, aligned sequences length are equal for both directions.",
      style="color:#d37217 ; font-weight: bold ;")
    DIRincorrect <- T # Direction determination is not sucessfull
    seqdirection_s2 <- "undetermined"
  }
}
```


```{r text direction determination s2, results='asis', eval=useSeqs2}
if(seqdirection_s2=="Forward") {
  if(TrimOKF==T) {
    cat("Sequencing result **#2** corresponds to **forward** direction")
    pa_F <- pa_s2_S
    seqtable_trim_F <- seqtable_trim_s2
  } else if (TrimOKF==F) {
    p("Warning: Insufficient quality of trimmed sequence for direction determination.",
      style="color:#d37217 ; font-weight: bold ;")
  }
}

if(seqdirection_s2=="Reverse") {
  if(TrimOKR==T) {
    cat("Sequencing result **#2** corresponds to **reverse** direction")
    pa_R <- pa_s2_RC
    seqtable_trim_R <- seqtable_trim_s2
  } else if (TrimOKR==F) {
    p("Warning: Insufficient quality of trimmed sequence for direction determination.",
      style="color:#d37217 ; font-weight: bold ;")
  }
}
```


```{r if same direction for the 2 sequencing, eval=useSeqs1 & useSeqs2}
if(seqdirection_s1==seqdirection_s2) {DIRincorrect <-T}
```


```{r output if no direction found, results='asis', eval=DIRincorrect}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if no direction found, eval=DIRincorrect}
# Diagram of directories
diagram_indiv
```

```{r stop knit if no direction found, results='asis', eval=DIRincorrect}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error: Analysis has been stopped as sequencing result direction (forward or reverse) could not be found.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1) 
knit_exit()
```





```{r do not use seq if trimming was not possible, include=F}
if(useSeqs1==T & seqdirection_s1=="Forward" & TrimOKs1==F) {useSeqF <- F}
if(useSeqs1==T & seqdirection_s1=="Reverse" & TrimOKs1==F) {useSeqR <- F}

if(useSeqs2==T & seqdirection_s2=="Forward" & TrimOKs2==F) {useSeqF <- F}
if(useSeqs2==T & seqdirection_s2=="Reverse" & TrimOKs2==F) {useSeqR <- F}
```


```{r get useSeq depending on direction for false values, include=F}
if(useSeqs1==F & seqdirection_s2=="Forward") {useSeqR <- F}
if(useSeqs1==F & seqdirection_s2=="Reverse") {useSeqF <- F}

if(useSeqs2==F & seqdirection_s1=="Forward") {useSeqR <- F}
if(useSeqs2==F & seqdirection_s1=="Reverse") {useSeqF <- F}

shiny::setProgress(0.5) 
```




```{r check if no sequencing exploitable, include=F}
# Verify if no CG covered by both sequencing
if (useSeqF==F & useSeqR==F) {NoSeq <- T} else {NoSeq <- F}
```
  
```{r output if no sequencing exploitable, results='asis', eval=NoSeq}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if no sequencing exploitable, eval=NoSeq}
# Diagram of directories
diagram_indiv
```

```{r stop knit if no sequencing exploitable, results='asis', eval=NoSeq}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error: Analysis has been stopped as none of the provided sequencing are of sufficient quality to be used.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1) 
knit_exit()
```




<!-- ALIGNED POSITIONS -->

```{r alignments positions F, include=F, eval=useSeqF}
# Get positions of start and end of the aligned sequences
alg_posF <- list(
  # F Subject : ref DNA BS
  subject_F_start = pa_F@subject@range@start,
  subject_F_end = pa_F@subject@range@start+pa_F@subject@range@width,
  # F Pattern : Sequencing results forward
  pattern_F_start = pa_F@pattern@range@start,
  pattern_F_end = pa_F@pattern@range@start+pa_F@pattern@range@width
)
```

```{r alignments positions R, include=F, eval=useSeqR}
# Get positions of start and end of the aligned sequences
alg_posR <- list(
  # R Subject : ref DNA BS RC
  subject_R_start = pa_R@subject@range@start,
  subject_R_end = pa_R@subject@range@start+pa_R@subject@range@width, 
  # R Pattern : Sequencing results reverse
  pattern_R_start = pa_R@pattern@range@start,
  pattern_R_end = pa_R@pattern@range@start+pa_R@pattern@range@width 
)
```

```{r alignments positions, include=F}
if (useSeqF==T & useSeqR==T) {alg_pos <- c(alg_posF,alg_posR)}
if (useSeqF==T & useSeqR==F) {alg_pos <- alg_posF}
if (useSeqF==F & useSeqR==T) {alg_pos <- alg_posR}
```


<!-- ALIGNED COORDINATES -->


```{r aligned sequences coordinates F, include=F, eval=useSeqF}
# Find coordinates of the aligned sequences on genome :
if(seq_strand=="plus") {
  alg_coordF = list(
    F_start = seq_start+alg_pos$subject_F_start-1, 
    F_end = seq_start+alg_pos$subject_F_end-1 ) 
}

if(seq_strand=="minus") {
  alg_coordF = list(
    F_start = seq_end-alg_pos$subject_F_end+1 , 
    F_end = seq_end-alg_pos$subject_F_start+1 ) 
}
```

```{r aligned sequences coordinates R, include=F, eval=useSeqR}
# Find coordinates of the aligned sequences on genome :
if(seq_strand=="plus") {
  alg_coordR = list(
    R_start = seq_end-alg_pos$subject_R_end+1 , 
    R_end = seq_end-alg_pos$subject_R_start+1) 
} 
if(seq_strand=="minus") {
  alg_coordR = list(
    R_start = seq_start+alg_pos$subject_R_start-1 , 
    R_end = seq_start+alg_pos$subject_R_end-1) 
}
```

```{r aligned sequences coordinates, include=F}
if (useSeqF==T & useSeqR==T) {alg_coord <- c(alg_coordF,alg_coordR)}
if (useSeqF==T & useSeqR==F) {alg_coord <- alg_coordF}
if (useSeqF==F & useSeqR==T) {alg_coord <- alg_coordR}
```







### Forward {.tabset .tabset-pills}

To find nucleotides of reference DNA sequence positions on sequencing result, the trimmed sequence of primary peaks from **forward** sequencing results is aligned with bisulfite converted reference DNA **sense** sequence.


#### Alignment

```{r warning if no aligment F,results='asis'}
if (TrimOKF==F) {
  p("Warning: Insufficient length or quality of trimmed forward sequencing.",
    style="color:#d37217 ; font-weight: bold ;")
}
```

```{r display alignment F, eval=useSeqF}
# Save alignment in .pair format (open in browser)
pa_F_path <- file.path(indivDir$alg,paste0(fileprefix,"_alignment_F.pair"))
writePairwiseAlignments(pa_F, file=pa_F_path, block.width=20) 
# display alignment
writePairwiseAlignments(pa_F,block.width=20)
```






#### Aligned sequences

```{r warning if no aligned seq F,results='asis'}
if (TrimOKF==F) {
  p("Warning: Insufficient length or quality of trimmed forward sequencing.",
    style="color:#d37217 ; font-weight: bold ;")
}
```


```{r aligned Patt F length, results='asis', eval=useSeqF}
p("Aligned sequence from forward sequencing result (Pattern P1):", style="font-weight: bold ;")
p("Length: ",nchar(as.character(alignedPattern(pa_F)))," bp") 
```

```{r aligned Patt F, eval=useSeqF}
cat(as.character(alignedPattern(pa_F)))
fasta_name <- paste(seq_name, 
                    "aligned sequence from forward sequencing result", 
                    paste0(nchar(alignedPattern(pa_F)),"bp"),
                    paste0(seq_chr, ":", alg_coord$F_start, "-", alg_coord$F_end),
                    sep=" ")
write.fasta(as.vector(alignedPattern(pa_F)), names=fasta_name, 
            file.out = file.path(indivDir$alg,paste0(fileprefix,"_aligned_seqresult_F.fasta")))
rm(fasta_name)
```



```{r aligned Sub F length, results='asis', eval=useSeqF}
p("Aligned sequence from bisulfite converted reference DNA sense sequence (Subject S1):", style="font-weight: bold ;")
p("Length: ",nchar(as.character(alignedSubject(pa_F)))," bp")
```

```{r aligned Sub F, eval=useSeqF}
cat(as.character(alignedSubject(pa_F)))
fasta_name <- paste(seq_name, 
                    "aligned sequence from bisulfite converted reference DNA sense sequence", 
                    paste0(nchar(alignedSubject(pa_F)),"bp"),
                    paste0(seq_chr, ":", alg_coord$F_start, "-", alg_coord$F_end),
                    sep=" ")
write.fasta(as.vector(alignedSubject(pa_F)), names=fasta_name, 
            file.out = file.path(indivDir$alg,paste0(fileprefix,"_aligned_DNAref_F.fasta")))
rm(fasta_name)
```




### Reverse {.tabset .tabset-pills}

To find nucleotides of reference DNA sequence positions on sequencing result, the trimmed sequence of primary peaks from **reverse** sequencing results is aligned with bisulfite converted reference DNA **antisense** sequence. 

#### Alignment

```{r warning if no aligment R,results='asis'}
if (TrimOKR==F) {
  p("Warning: Insufficient length or quality of trimmed reverse sequencing.",
    style="color:#d37217 ; font-weight: bold ;")
}
```

```{r display aligment R, eval=useSeqR}
# Save alignments in .pair format (open in browser)
pa_R_path <- file.path(indivDir$alg,paste0(fileprefix,"_alignment_R.pair"))
writePairwiseAlignments(pa_R, file=pa_R_path, block.width=20) 
# display alignments
writePairwiseAlignments(pa_R,block.width=20)
```



#### Aligned sequences

```{r warning if no aligned seq R,results='asis'}
if (TrimOKR==F) {
  p("Warning: Insufficient length or quality of trimmed reverse sequencing.",
    style="color:#d37217 ; font-weight: bold ;")
}
```



```{r aligned Patt R length, results='asis', eval=useSeqR}
p("Aligned sequence from reverse sequencing result (Pattern P1):", style="font-weight: bold ;")
p("Length: ",nchar(as.character(alignedPattern(pa_R)))," bp")
```

```{r aligned Patt R, eval=useSeqR}
cat(as.character(alignedPattern(pa_R)))
fasta_name <- paste(seq_name, 
                    "aligned sequence from reverse sequencing result", 
                    paste0(nchar(alignedPattern(pa_R)),"bp"),
                    paste0(seq_chr, ":", alg_coord$R_start, "-", alg_coord$R_end),
                    sep=" ")
write.fasta(as.vector(alignedPattern(pa_R)), names=fasta_name, 
            file.out = file.path(indivDir$alg,paste0(fileprefix,"_aligned_seqresult_R.fasta")))
rm(fasta_name)
```



```{r aligned Sub R length, results='asis', eval=useSeqR}
p("Aligned sequence from bisulfite converted reference DNA antisense sequence (Subject S1):", style="font-weight: bold ;")
p("Length: ",nchar(as.character(alignedSubject(pa_R)))," bp")
```

```{r aligned Sub R, eval=useSeqR}
cat(as.character(alignedSubject(pa_R)))
fasta_name <- paste(seq_name, 
                    "aligned sequence from bisulfite converted reference DNA antisense sequence", 
                    paste0(nchar(alignedSubject(pa_R)),"bp"),
                    paste0(seq_chr, ":", alg_coord$R_start, "-", alg_coord$R_end),
                    sep=" ")
write.fasta(as.vector(alignedSubject(pa_R)), names=fasta_name, 
            file.out = file.path(indivDir$alg,paste0(fileprefix,"_aligned_DNAref_R.fasta")))
rm(fasta_name)
```







## Quality control {.tabset .tabset-pills}

```{r QC Thresholds}
# The aim of the quality control is to control the sequencing quality regarding the reference sequence, using alignments

# Thresholds 
th_identity <- 75
th_conv_rate <- 0.90

indel <- list()

shiny::setProgress(0.6) 
```

```{r QC positions of gaps F, eval=useSeqF}
# retrieve positions of gaps :

## On ref DNA sequence (Subject):
indel$dna_ref_F <- alg_pos$subject_F_start + pa_F@subject@indel@unlistData@start -1
## On sequencing (Pattern) :
indel$sequencing_F = alg_pos$pattern_F_start + pa_F@pattern@indel@unlistData@start -1
```

```{r QC positions of gaps R, eval=useSeqR}
# retrieve positions of gaps :
 
## On ref DNA sequence (Subject):
indel$dna_ref_R <- alg_pos$subject_R_start + pa_R@subject@indel@unlistData@start -1
## On sequencing (Pattern) :
indel$sequencing_R = alg_pos$pattern_R_start + pa_R@pattern@indel@unlistData@start -1
```

```{r QC bisulfite conv rate}
# Get bisulfite conversion rate
BisConRate <- BisulfiteConvRate(sequences=seq, alg_pos=alg_pos, seqtable_trim_F=seqtable_trim_F, seqtable_trim_R=seqtable_trim_R,
                                indel = indel, useSeqF = useSeqF, useSeqR = useSeqR)

if (useSeqF==T & useSeqR==T) {
  # Join two data frame to have all data from both sequencing
  BisConRate_all <- full_join(BisConRate$F[,c(1,9)],BisConRate$R[,c(1,9)], by="C_nb", suffix = c("_F", "_R"))
}
if (useSeqF==T & useSeqR==F) {
  BisConRate_all <- BisConRate$F[,c(1,9)]
}
if (useSeqF==F & useSeqR==T) {
  BisConRate_all <- BisConRate$R[,c(1,9)]
}
```

```{r QC table row F, include=F, eval=useSeqF}
QC_table_F <- data.frame(
  check.names = FALSE,
  "Direction" = "Forward",
  "Sequencing" = paste0("#",seqnumber_F),
  "Length" = nchar(subject(pa_F)), 
  "Identity" = round(pid(pa_F),2),
  "Mismatch" = nmismatch(pa_F),
  "Indel in DNA reference" = pa_F@subject@indel@partitioning@end, 
  "Indel in sequencing" = pa_F@pattern@indel@partitioning@end, 
  "Mean of conversion rates" = round(mean(BisConRate$F$conversion_rate, na.rm=T),2),
  "SD of conversion rates" = round(sd(BisConRate$F$conversion_rate, na.rm=T),2)
)

if (QC_table_F[1, "Length"] >= th_min_alg &
    QC_table_F[1, "Identity"] >= th_identity &
    QC_table_F[1, "Mean of conversion rates"] >= th_conv_rate) {
  # Quality of forward sequencing is good enougth :
  correct_F <- T
} else {
  # Quality of forward sequencing is not good enougth :
  correct_F <- F
}
if (useSeqR==F) {correct_R<-F}

QC_table_F[, "Quality Result"] = ifelse(correct_F==T,"Correct","Incorrect")
```

```{r QC table row R, include=F, eval=useSeqR}
QC_table_R <- data.frame(
  check.names = FALSE,
  "Direction" = "Reverse",
  "Sequencing" = paste0("#",seqnumber_R),
  "Length" = nchar(subject(pa_R)),
  "Identity" = round(pid(pa_R),2), 
  "Mismatch" = nmismatch(pa_R), 
  "Indel in DNA reference" = pa_R@subject@indel@partitioning@end, 
  "Indel in sequencing" = pa_R@pattern@indel@partitioning@end, 
  "Mean of conversion rates" = round(mean(BisConRate$R$conversion_rate, na.rm=T),2),
  "SD of conversion rates" = round(sd(BisConRate$R$conversion_rate, na.rm=T),2)
)

if (QC_table_R[1, "Length"] >= th_min_alg &
    QC_table_R[1, "Identity"] >= th_identity &
    QC_table_R[1, "Mean of conversion rates"] >= th_conv_rate) {
  # Quality of forward sequencing is good enougth :
  correct_R <- T
} else {
  # Quality of forward sequencing is not good enougth :
  correct_R <- F
}
if (useSeqF==F) {correct_F<-F}

QC_table_R[, "Quality Result"] = ifelse(correct_R==T,"Correct","Incorrect")
```

```{r QC create table}
# Create table for quality control visualization

if (useSeqF==T & useSeqR==T) {QC_table <- rbind(QC_table_F, QC_table_R )}
if (useSeqF==T & useSeqR==F) {QC_table <- QC_table_F}
if (useSeqF==F & useSeqR==T) {QC_table <- QC_table_R}
```

### Summary

| Thresholds                                  | Values                   |
|:-------------|:--------------------------|
| Minimum length of the aligned sequence :        | **`r th_min_alg` bp**    |
| Minimum identity percentage of alignment :  | **`r th_identity` %**    | 
| Minimum of bisulfite conversion rate mean : | **`r th_conv_rate`**     |
  
  
```{r QC display table, results='asis'}
# Export table data as csv and xlsx
write.table(QC_table, file = file.path(indivDir$tab,paste0(fileprefix,"_QC_table.csv")), row.names = T, col.names = T)
write.xlsx(QC_table, file = file.path(indivDir$tab,paste0(fileprefix,"_QC_table.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display Quality table
QC_table_FT <- formattable(
  QC_table, 
  align = c("l", rep("c",ncol(QC_table)-1)),
  list("Direction" = formatter("span", style = ~ formattable::style(font.weight = "bold")),
       "Sequencing" = formatter("span", style = ~ formattable::style(font.weight = "bold")),
       "Length" = color_condi(">=", th_min_alg, colors_tables["correct_pale"], colors_tables["incorrect_pale"]),
       "Identity" = color_condi(">=", th_identity, colors_tables["correct_pale"], colors_tables["incorrect_pale"]),
       "Mismatch" = color_condi("<=", 20, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Indel in DNA reference" = color_condi("<", 1, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Indel in sequencing" = color_condi("<", 1, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Mean of conversion rates" = color_condi(">=", th_conv_rate,colors_tables["correct_pale"], colors_tables["incorrect_pale"]),
       "SD of conversion rates" = color_condi("<=", 0.25, colors_tables["correct_pale"], colors_tables["warning_pale"]),
       "Quality Result" = formatter("span", x ~ icontext(ifelse(x=="Correct", "ok", "remove"), ifelse(x=="Correct", "Correct", "Incorrect")), 
                                    style = x ~ formattable::style(color = ifelse(x=="Correct", colors_tables["correct"], colors_tables["incorrect"])))))

QC_table_FT

if(TrimOKF==F) { 
  p("Forward sequencing result could have not been trimmed due to poor quality.",
  style="color:#bf3232 ; font-weight: bold ;")
}
if(TrimOKR==F) { 
  p("Reverse sequencing result could have not been trimmed due to poor quality.",
    style="color:#bf3232 ; font-weight: bold ;")
}

if("Incorrect" %in% QC_table[,"Quality Result"]) {
  p("Warning: Incorrect sequencing results are excluded from methylation quantification.",
  style="color:#d37217 ; font-weight: bold ;")
}
if(length(QC_table[,"Indel in DNA reference"]==0) != length(QC_table[,"Indel in DNA reference"]) |
   length(QC_table[,"Indel in sequencing"]==0) != length(QC_table[,"Indel in sequencing"])) {
  p("Warning: Gaps (insertion/deletions) in alignments are taken into account to retrieve correct C positions and associated peak base values.",
    style="color:#d37217 ; font-weight: bold ;")
}
```

```{r QC export table, include=F}
export_formattable(QC_table_FT,file=file.path(indivDir$tab,paste0(fileprefix,"_QC.png")))
```



  
### Mismatch positions {.tabset .tabset-pills} 

#### Forward
```{r QC mismatch F, results='asis', eval=useSeqF}
mismatch_F <- mismatchTable(pa_F)
if (nrow(mismatch_F)>0) {
QC_mismatch_F <- data.frame(
  "Positions in trimmed sequencing result" = mismatch_F$PatternStart,
  "Nucleotides in trimmed sequencing result" = mismatch_F$PatternSubstring,
  "Positions in reference DNA sequence"= mismatch_F$SubjectStart,
  "Nucleotides in reference DNA sequence" = mismatch_F$SubjectSubstring, check.names=F)
formattable(QC_mismatch_F, align=rep("c",ncol(QC_mismatch_F)))} else {
  cat("No mismatch positions in forward alignment.")
}
```



#### Reverse
```{r QC mismatch R, results='asis', eval=useSeqR}
mismatch_R <- mismatchTable(pa_R)
if (nrow(mismatch_R)>0) {
QC_mismatch_R <- data.frame(
  "Positions in trimmed sequencing result" = mismatch_R$PatternStart,
  "Nucleotides in trimmed sequencing result" = mismatch_R$PatternSubstring,
  "Positions in reference DNA sequence"= mismatch_R$SubjectStart,
  "Nucleotides in reference DNA sequence" = mismatch_R$SubjectSubstring, check.names=F)
formattable(QC_mismatch_R, align=rep("c",ncol(QC_mismatch_R)))} else {
  cat("No mismatch positions in reverse alignment.")
}
```



### Insertions/deletions {.tabset .tabset-pills}

#### Forward
```{r QC indel F 2, results='asis', eval=useSeqF}
if (pa_F@pattern@indel@partitioning@end!=0 | pa_F@subject@indel@partitioning@end!=0) {
  data_length <- max(length(indel$sequencing_F),length(indel$dna_ref_F))
  QC_indel_F <- data.frame(matrix(nrow=data_length,ncol=2))
  colnames(QC_indel_F) <- c("Indel positions in \n trimmed sequencing result","Indel positions in \n reference DNA sequence")
  QC_indel_F[,1] <- indel$sequencing_F
  QC_indel_F[,2] <- indel$dna_ref_F
  QC_indel_F[which(is.na(QC_indel_F)==T)] <- ""
  formattable(QC_indel_F, align=c("c","c"))} else {
  cat("No insertions/deletions in forward alignment.")
}
```



#### Reverse
```{r QC indel R 2, results='asis', eval=useSeqR}
if (pa_R@pattern@indel@partitioning@end!=0 | pa_R@subject@indel@partitioning@end!=0) {
  data_length <- max(length(indel$sequencing_R),length(indel$dna_ref_R))
  QC_indel_R <- data.frame(matrix(nrow=data_length,ncol=2))
  colnames(QC_indel_R) <- c("Indel positions in \n sequence from sequencing result","Indel positions in \n reference DNA sequence")
  QC_indel_R[,1] <- indel$sequencing_R
  QC_indel_R[,2] <- indel$dna_ref_R
  QC_indel_R[which(is.na(QC_indel_R)==T)] <- ""
  formattable(QC_indel_R, align=c("c","c"))} else {
  cat("No insertions/deletions in reverse alignment.")
}
```




### Conversion rates  {.tabset .tabset-pills}

#### Combined
```{r QC Conversion rates combined}
# Export table data as csv and xlsx
write.table(BisConRate_all, file = file.path(indivDir$tab,paste0(fileprefix,"_bisconv_table.csv")), row.names = T, col.names = T)
write.xlsx(BisConRate_all, file = file.path(indivDir$tab,paste0(fileprefix,"_bisconv_table.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display
if (useSeqF==T & useSeqR==T) {
  colnames(BisConRate_all) <- c("C number","Bisulfite conversion rate Forward","Bisulfite conversion rate Reverse")
  BisConRate_all[,c(2,3)] <- format(round(BisConRate_all[,c(2,3)],2),nsmall=2)
  BisConRate_all_FT <- formattable(BisConRate_all,
                                   align = rep("c",ncol(BisConRate_all)),
                                   width = "50%",
                                   list("Bisulfite conversion rate Forward" = color_condi(">=", 
                                                                                          th_conv_rate,colors_tables["correct_pale"], colors_tables["incorrect_pale"]),
                                        "Bisulfite conversion rate Reverse" = color_condi(">=", 
                                                                                          th_conv_rate,colors_tables["correct_pale"], colors_tables["incorrect_pale"])))
}

if (useSeqF==T & useSeqR==F) {
  colnames(BisConRate_all) <- c("C number","Bisulfite conversion rate Forward")
  BisConRate_all[,2] <- format(round(BisConRate_all[,2],2),nsmall=2)
  BisConRate_all_FT <- formattable(BisConRate_all,
                                   align = rep("c",ncol(BisConRate_all)),
                                   width = "50%",
                                   list("Bisulfite conversion rate Forward" = color_condi(">=", 
                                                                                          th_conv_rate,colors_tables["correct_pale"], colors_tables["incorrect_pale"]))) 
}

if (useSeqF==F & useSeqR==T) {
  colnames(BisConRate_all) <- c("C number","Bisulfite conversion rate Reverse")
  BisConRate_all[,2] <- format(round(BisConRate_all[,2],2),nsmall=2)
  BisConRate_all_FT <- formattable(BisConRate_all,
                                   align = rep("c",ncol(BisConRate_all)),
                                   width = "50%",
                                   list("Bisulfite conversion rate Reverse" = color_condi(">=", 
                                                                                          th_conv_rate,colors_tables["correct_pale"], colors_tables["incorrect_pale"]))) 
}

BisConRate_all_FT
```


#### Forward
```{r QC Conversion rates F, eval=useSeqF}
colnames(BisConRate$F) <- c("C number","Position in reference DNA", "Position in sequencing result (trimmed)","Position in sequencing result (raw)",
                            "A peak", "T peak", "G peak", "C peak", "Bisulfite conversion rate")

formattable(BisConRate$F,
            align=rep("c",ncol(BisConRate$F)),
            list("T peak"= color_format("white",bases_colors["T"]),
                 "C peak"= color_format("white",bases_colors["C"]),
                 "Bisulfite conversion rate" = color_condi(">=", th_conv_rate, colors_tables["correct_pale"], colors_tables["incorrect_pale"])))
```



#### Reverse
```{r QC Conversion rates R, eval=useSeqR}
colnames(BisConRate$R) <- c("C number","Position in reference DNA", "Position in sequencing result (trimmed)", "Position in sequencing result (raw)",
                            "A peak", "T peak", "G peak", "C peak", "Bisulfite conversion rate")

formattable(BisConRate$R,
            align=rep("c",ncol(BisConRate$R)),
            list("A peak"= color_format("white",bases_colors["A"]),
                 "G peak"= color_format("white",bases_colors["G"]),
                 "Bisulfite conversion rate" = color_condi(">=", th_conv_rate, colors_tables["correct_pale"], colors_tables["incorrect_pale"])))
```




```{r QC Conversion rates combined export table, include=F}
export_formattable(BisConRate_all_FT,file=file.path(indivDir$tab,paste0(fileprefix,"_bisconv_table.png")))
shiny::setProgress(0.7)
```



```{r check if QC incorrect}
# Verify if both sequencing are defined incrorrect by QC
if (useSeqF==T & useSeqR==T & correct_F==F & correct_R==F) { QCincorrect <- T 
} else if (useSeqF==T & useSeqR==F & correct_F==F) { QCincorrect <- T 
} else if (useSeqF==F & useSeqR==T & correct_R==F) { QCincorrect <- T 
} else { QCincorrect <- F}
```
  
```{r output if QC incorrect, results='asis', eval=QCincorrect}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if QC incorrect, eval=QCincorrect}
# Diagram of directories
diagram_indiv
```

```{r stop knit if QC incorrect, results='asis', eval=QCincorrect}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error: Analysis has been stopped as sequencing results are defined as incorrect by Quality Control. The analysis can not be performed.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1) 
knit_exit()
```



### Maximum aligned sequence  

The maximum aligned sequence corresponds to the maximum sequence covered by at least one correct sequencing.

```{r max aligned sequence coordinates, results='asis'}
# Larger coordinates of sequence covered by at least one of the two sequencing (called max aligned sequence)

# Both sequencing
if (useSeqF==T & useSeqR==T) {
  if(correct_F==T & correct_R==T) {
    alg_coord <- c(alg_coord, 
                   min_start = min(alg_coord$F_start,alg_coord$R_start),
                   max_end = max(alg_coord$F_end,alg_coord$R_end))
  }
  if(correct_F==T & correct_R==F) {
    alg_coord <- c(alg_coord,
                   min_start = alg_coord$F_start,
                   max_end = alg_coord$F_end) 
    p("Warning: Incorrect reverse sequencing, positions of maximum aligned sequence are taken from forward sequencing aligned sequence.",
      style="color:#d37217 ; font-weight: bold ;")
  }
  
  if(correct_F==F & correct_R==T) {
    alg_coord <- c(alg_coord,
                   min_start = alg_coord$R_start,
                   max_end = alg_coord$R_end)
    p("Warning: Incorrect forward sequencing, positions of maximum aligned sequence are taken from forward sequencing aligned sequence.",
      style="color:#d37217 ; font-weight: bold ;")
  }
}

# Forward
if (useSeqF==T & useSeqR==F) {
  alg_coord <- c(alg_coord,
                 min_start = alg_coord$F_start,
                 max_end = alg_coord$F_end) 
}

if (useSeqF==F & useSeqR==T) {
  alg_coord <- c(alg_coord,
                 min_start = alg_coord$R_start,
                 max_end = alg_coord$R_end)
}

```

```{r get max aligned sequence, include=F}
# Get positions of max aligned sequence on reference DNA sequence to extract its sequence
seq_max_aligned <- getMaxAligned(sequenceBS = seq_BS$seq_BS, seq_strand = seq_strand, 
                                 alg_coord_start = alg_coord$min_start, alg_coord_end = alg_coord$max_end,
                                 seq_start = seq_start, seq_end = seq_end)
alg_coord <- c(alg_coord, max_length = nchar(as.character(seq_max_aligned)))

# Save max aligned sequence as fasta
fasta_name <- paste(seq_name, 
                    "maximum aligned sequence from both sequencing", 
                    paste0(alg_coord$max_length,"bp"),
                    paste0(seq_chr, ":", alg_coord$min_start, "-", alg_coord$max_end),
                    sep=" ")
write.fasta(as.vector(seq_max_aligned), names=fasta_name, 
            file.out = file.path(indivDir$alg,paste0(fileprefix,"_aligned_max_seq.fasta")))
rm(fasta_name)
```

Length of max aligned sequence : **`r paste0(as.character(alg_coord$max_length)," bp")`**  
Coordinates of max aligned sequence : **`r paste0(seq_chr,":",alg_coord$min_start,"-",alg_coord$max_end)`**
  
**Maximum aligned sequence** (on bisulfite converted sense sequence from the `r seq_strand` strand sequence) :
```{r display max aligned sequence}
cat(as.character(seq_max_aligned))

shiny::setProgress(0.8) 
```



```{r check if no CG covered}
# Verify if no CG covered by both sequencing
if (useSeqF==T & useSeqR==T & correct_F==T & correct_R==T) {
  NoCGcovered_F <- checkCGcover(data_pos=CG_pos$CG_pos, subject_start=alg_pos$subject_F_start, subject_end=alg_pos$subject_F_end)
  NoCGcovered_R <- checkCGcover(data_pos=CG_pos$CG_pos_RC, type="RC", subject_start=alg_pos$subject_R_start, subject_end=alg_pos$subject_R_end)
  if (NoCGcovered_F==T & NoCGcovered_R==T) {NoCGcovered <- T} else {NoCGcovered <- F}
}
if (useSeqF==T & useSeqR==F | correct_F==T & correct_R==F ) {
  NoCGcovered <- checkCGcover(data_pos=CG_pos$CG_pos, subject_start=alg_pos$subject_F_start, subject_end=alg_pos$subject_F_end)
}
if (useSeqF==F & useSeqR==T | correct_F==F & correct_R==T ) {
  NoCGcovered <- checkCGcover(data_pos=CG_pos$CG_pos_RC, type="RC", subject_start=alg_pos$subject_R_start, subject_end=alg_pos$subject_R_end)
}
```
  
```{r output if no CG covered, results='asis', eval=NoCGcovered}
# Output tab and Directories tab
cat("## Output data {.tabset .tabset-pills}","### Directories", sep="  \n")
```

```{r diagram if no CG covered, eval=NoCGcovered}
# Diagram of directories
diagram_indiv
```

```{r stop knit if no CG covered, results='asis', eval=NoCGcovered}
# Files tab with list of output files
cat("### Files","  \n")
list_output_indiv(fileprefix = fileprefix)
# Abort analysis and error message
cat("## Analysis aborted","  \n")
p("Error: Analysis has been stopped as no CpG sites were found covered by sequencing results.",
  style="color:#bf3232 ; font-weight: bold ;")
shiny::setProgress(1) 
knit_exit()
```


## Methylation {.tabset .tabset-pills}


```{r}
if (useSeqF==T) { CG_peaks_F <- getPeaks(CG_pos$CG_pos, alg_pos = alg_pos, seqtable_trim = seqtable_trim_F, indel = indel, sequencing = "forward") }
if (useSeqR==T) { CG_peaks_R <- getPeaks(CG_pos$CG_pos_RC, alg_pos = alg_pos, seqtable_trim = seqtable_trim_R, indel = indel, sequencing = "reverse") }
```


### Combined  

```{r combine F and R methylation results, results='asis', warning=F}

# Combine dataframes from Forward sequencing and Reverse Sequencing
if (useSeqF==T & useSeqR==T & correct_F==T & correct_R==T) {
  CG_meth <- full_join(subset(CG_peaks_F, select= c(CG_nb,chr,start,end,meth_F)),
                       subset(CG_peaks_R, select= c(CG_nb,chr,start,end,meth_R)),
                       by = c("CG_nb","chr","start","end"))
  # Order by CG_nb 
  CG_meth = CG_meth %>% 
    .[with(., order(CG_nb)),]
  # Reset row number
  rownames(CG_meth) <- NULL 
  # Add mean and sd of methylation
  CG_meth = CG_meth %>% 
    rowwise() %>% 
    dplyr::mutate(meth_avg=mean(c(meth_F,meth_R),na.rm=T)) %>%
    dplyr::mutate(meth_sd=round(sd(c(meth_F,meth_R)),2))
  
  p("Methylation mean of entire sequence (CpG covered) : ", round(mean(CG_meth$meth_avg),2)," %", style="font-weight: bold ;")
  
  # Preparation for display
  CG_meth_FT <- CG_meth
  colnames(CG_meth_FT) <- c("CG number","Chromosome","Start","End","Methylation Percentage Forward", "Methylation Percentage Reverse",
                            "Methylation Percentage Average", "Methylation Percentage SD")
  CG_meth_FT <- formattable(CG_meth_FT,
                            align=rep("c",ncol(CG_meth_FT)),
                            list("Methylation Percentage Forward" = color_meth(),
                                 "Methylation Percentage Reverse" = color_meth(),
                                 "Methylation Percentage Average" = color_meth(),
                                 "Methylation Percentage SD"= color_condi("<",25,colors_tables["correct_pale"], colors_tables["warning_pale"])))
}

if (useSeqF==T & useSeqR==F | correct_F==T & correct_R==F) {
  CG_meth <- subset(CG_peaks_F, select= c(CG_nb,chr,start,end,meth_F))
  CG_meth[,"meth_R"] <- NA
  CG_meth[,"meth_avg"] <- CG_meth$meth_F
  CG_meth[,"meth_sd"] <- NA
  p("Warning: Only forward sequencing data available.", style="color:#d37217 ; font-weight: bold ;")
  p("Methylation mean of entire sequence (CpG covered) : ", round(mean(CG_meth$meth_F),2)," %", style="font-weight: bold ;")
  # Preparation for display
  CG_meth_FT <- CG_meth
  colnames(CG_meth_FT) <- c("CG number","Chromosome","Start","End","Methylation Percentage Forward", "Methylation Percentage Reverse",
                            "Methylation Percentage Average", "Methylation Percentage SD")
  CG_meth_FT <- formattable(CG_meth_FT,
                            align=rep("c",ncol(CG_meth_FT)),
                            list("Methylation Percentage Forward" = color_meth(),
                                 "Methylation Percentage Reverse" = color_meth(),
                                 "Methylation Percentage Average" = color_meth(),
                                 "Methylation Percentage SD"= color_condi("<",25,colors_tables["correct_pale"], colors_tables["warning_pale"])))
}

if (useSeqF==F & useSeqR==T | correct_F==F & correct_R==T) {
  CG_meth <- subset(CG_peaks_R, select= c(CG_nb,chr,start,end,meth_R))
  CG_meth[,"meth_F"] <- NA
  CG_meth[,"meth_avg"] <- CG_meth$meth_R
  CG_meth <- CG_meth[,c("CG_nb","chr","start","end","meth_F","meth_R","meth_avg")]
  CG_meth[,"meth_sd"] <- NA
  p("Warning: Only reverse sequencing data available.", style="color:#d37217 ; font-weight: bold ;")
  p("Methylation mean of entire sequence (CpG covered) : ", round(mean(CG_meth$meth_R),2)," %", style="font-weight: bold ;")
  # Preparation for display
  CG_meth_FT <- CG_meth
  colnames(CG_meth_FT) <- c("CG number","Chromosome","Start","End","Methylation Percentage Forward", "Methylation Percentage Reverse",
                            "Methylation Percentage Average", "Methylation Percentage SD")
  CG_meth_FT <- formattable(CG_meth_FT,
                            align=rep("c",ncol(CG_meth_FT)),
                            list("Methylation Percentage Forward" = color_meth(),
                                 "Methylation Percentage Reverse" = color_meth(),
                                 "Methylation Percentage Average" = color_meth(),
                                 "Methylation Percentage SD"= color_condi("<",25,colors_tables["correct_pale"], colors_tables["warning_pale"])))
}
  
  
# Export table data as csv and xlsx
write.table(CG_meth, file = file.path(indivDir$tab,paste0(fileprefix,"_meth_table.csv")), row.names = T, col.names = T)
write.xlsx(CG_meth, file = file.path(indivDir$tab,paste0(fileprefix,"_meth_table.xlsx")), rowNames = T, colNames = T, overwrite = T)

# Display table
CG_meth_FT
```

```{r meth export table, include=F, warning=F}
export_formattable(CG_meth_FT,file=file.path(indivDir$tab,paste0(fileprefix,"_meth_table.png")))
```



### Forward  

```{r peaks meth F, results='asis'}
# Forward sequening methylation result

# No forward sequencing provided :
if (useSeqF==F) { p("Warning: No sequencing data from foward direction sequencing.", style="color:#d37217 ; font-weight: bold ;")}


# Forward Sequening provided :
if (useSeqF==T) {
  
  # Incorrect Forward Sequencing
  if (correct_F==F) { p("Warning : Quality Control has defined the forward sequencing as incorrect.", style="color:#d37217 ; font-weight: bold ;")}
  
  # Correct Forward Sequencing
  if (correct_F==T) {
    
    # Display Methylation dataframe for Forward Sequencing
    if (nrow(CG_peaks_F)!=0) {
      CG_peaks_F_FT <- CG_peaks_F[,-6]
      colnames(CG_peaks_F_FT) <- c("CG number","Chromosome","Start","End","C position in reference DNA", 
                                   "C position in sequencing result (trimmed)", "C position in sequencing result (raw)",
                                   "A peak", "T peak", "G peak", "C peak", "Methylation Percentage")
      formattable(CG_peaks_F_FT,
                  align=rep("c",ncol(CG_peaks_F_FT)),
                  list("T peak"= color_format("white",bases_colors["T"]),
                       "C peak"= color_format("white",bases_colors["C"]),
                       "Methylation Percentage" = color_meth()))
      
    } else {
      p("Warning: Table with 0 rows, 0 CG covered by sequencing from forward primer.", style="color:#d37217 ; font-weight: bold ;")
    }
  }
}

```





### Reverse  

```{r peaks meth R, results='asis'}
# Reverse sequening methylation result

# No reverse sequencing provided :
if (useSeqR==F) { p("Warning: No sequencing data from reverse direction sequencing.", style="color:#d37217 ; font-weight: bold ;")}


# Reverse sequening provided :
if (useSeqR==T) {
  
  # Incorrect Reverse Sequencing
  if (correct_R==F) { p("Warning : Quality Control has defined the reverse sequencing as incorrect.", style="color:#d37217 ; font-weight: bold ;")}
  
  # Correct Reverse Sequencing
  if (correct_R==T) {
    
    # Display Methylation dataframe for Reverse Sequencing
    if (nrow(CG_peaks_R)!=0) {
      CG_peaks_R_FT <- CG_peaks_R[,-5]
      colnames(CG_peaks_R_FT) <- c("CG number","Chromosome","Start","End","G position in reference DNA",
                                    "G position in sequencing result (trimmed)", "G position in sequencing result (raw)",
                                   "A peak", "T peak", "G peak", "C peak", "Methylation Percentage")
      formattable(CG_peaks_R_FT,
                  align=rep("c",ncol(CG_peaks_R_FT)),
                  list("A peak"= color_format("white",bases_colors["A"]),
                       "G peak"= color_format("white",bases_colors["G"]),
                       "Methylation Percentage" = color_meth()))
      
    } else {
      p("Warning: Table with 0 rows, 0 CG covered by sequencing from reverse primer.", style="color:#d37217 ; font-weight: bold ;")
    }
  }
}
```








```{r save meth data file, include=F, warning=F}
# Save methylation data files for grouped analysis
CG_meth_data = CG_meth %>% 
  dplyr::mutate(collection=seq_coll, group=seq_group, rep=seq_rep, clone=clone_nb,
                alg_coord_start=alg_coord$min_start, alg_coord_end=alg_coord$max_end)
colnames(CG_meth_data)[colnames(CG_meth_data)=="chr"] <- "chromosome"


data_path <- file.path(indivDir$data,paste0(fileprefix,"_methdata.csv"))
write.table(CG_meth_data, file = data_path, row.names = F)

shiny::setProgress(0.9) 
```





### Individual methylation plot

```{r plot preparation, include=F}
# Create variable for name of individual experiment
exp_name <- paste(seq_name,seq_coll,seq_group,seq_rep)
if(seq_clone!="" & !is.na(clone_nb)) {exp_name <- paste(exp_name,"clone",clone_nb)}
exp_name <- gsub("  "," ",exp_name)
exp_name <- gsub("* $","",exp_name)

# methylation data
CG_meth_plot <- CG_meth[,c(2,3,4,7)]
colnames(CG_meth_plot) <- c("chromosome","start","end",exp_name)
```

```{r plot legend, include=F}
# Generation of plot legend image if not already created
legend_bases_path <-  file.path(indivDir$plot,"legend_bases.png")

if (!file.exists(legend_bases_path)) { 
  png(filename=legend_bases_path, width=320, height=60)
  par(mai=c(0,0,0,0))
  plot.new()
  legend(title = "Colors of bases", 
         x="center", 
         legend=c("A","T","G","C","N"), 
         fill=bases_colors[1:5],
         border="white", 
         bty = "n", 
         cex=1.5,
         horiz=TRUE)
  dev.off()
}
```

```{r plot generation, include=F}
# Plot generation

plot_path <- file.path(indivDir$plot,paste0(fileprefix,"_plot.png"))
png(filename=plot_path, width = alg_coord$max_length*10, height = 600)
individual_meth_plot(meth=CG_meth_plot, start=alg_coord$min_start, end=alg_coord$max_end, genome = genome,
                     plot_title = paste("Methylation plot of",exp_name), bases_colors=bases_colors)
dev.off()
```


```{r display plot and legends, results='asis', message=F, warning=F, out.width = "100%", fig.align = "center"}
cat("\n\n![](",plot_path,")\n\n")
cat("\n\n![](",legend_bases_path,")")
```



## Output data {.tabset .tabset-pills}

### Directories
```{r display file chart, fig.align = "center"}
# Output folders chart
diagram_indiv
```


### Files
```{r list of output files, results='asis'}
# List of files related to analysis
list_output_indiv(fileprefix = fileprefix)
```


### Methylation data file preview
```{r meth data file, warning=F}
# Display methylation data file
formattable(CG_meth_data,
            align=rep("c",ncol(CG_meth_data)))

shiny::setProgress(1) 
```


<!--

> sessioninfo::session_info()
 Session info 
 setting  value
 version  R version 4.2.0 (2022-04-22 ucrt)
 os       Windows 10 x64 (build 19044)
 system   x86_64, mingw32
 ui       RStudio
 language (EN)
 collate  French_France.utf8
 ctype    French_France.utf8
 tz       Europe/Paris
 date     2022-05-18
 rstudio  2022.02.2+485 Prairie Trillium (desktop)
 pandoc   2.17.1.1 @ C:/Program Files/RStudio/bin/quarto/bin/ (via rmarkdown)

 Packages 
 package              * version  date (UTC) lib source
 abind                  1.4-5    2016-07-21 [1] CRAN (R 4.2.0)
 ade4                   1.7-19   2022-04-19 [1] CRAN (R 4.2.0)
 AnnotationDbi          1.58.0   2022-04-26 [1] Bioconductor
 AnnotationFilter       1.20.0   2022-04-26 [1] Bioconductor
 ape                  * 5.6-2    2022-03-02 [1] CRAN (R 4.2.0)
 arrangements         * 1.1.9    2020-09-13 [1] CRAN (R 4.2.0)
 askpass                1.1      2019-01-13 [1] CRAN (R 4.2.0)
 assertthat             0.2.1    2019-03-21 [1] CRAN (R 4.2.0)
 backports              1.4.1    2021-12-13 [1] CRAN (R 4.2.0)
 base64enc              0.1-3    2015-07-28 [1] CRAN (R 4.2.0)
 Biobase                2.56.0   2022-04-26 [1] Bioconductor
 BiocFileCache          2.4.0    2022-04-26 [1] Bioconductor
 BiocGenerics         * 0.42.0   2022-04-26 [1] Bioconductor
 BiocIO                 1.6.0    2022-04-26 [1] Bioconductor
 BiocManager          * 1.30.17  2022-04-22 [1] CRAN (R 4.2.0)
 BiocParallel           1.30.0   2022-04-27 [1] Bioconductor
 BiocStyle            * 2.24.0   2022-04-26 [1] Bioconductor
 biomaRt                2.52.0   2022-04-26 [1] Bioconductor
 Biostrings           * 2.64.0   2022-04-26 [1] Bioconductor
 biovizBase             1.44.0   2022-04-26 [1] Bioconductor
 bit                    4.0.4    2020-08-04 [1] CRAN (R 4.2.0)
 bit64                  4.0.5    2020-08-30 [1] CRAN (R 4.2.0)
 bitops                 1.0-7    2021-04-24 [1] CRAN (R 4.2.0)
 blob                   1.2.3    2022-04-10 [1] CRAN (R 4.2.0)
 broom                  0.8.0    2022-04-13 [1] CRAN (R 4.2.0)
 BSgenome             * 1.64.0   2022-04-26 [1] Bioconductor
 bslib                  0.3.1    2021-10-06 [1] CRAN (R 4.2.0)
 cachem                 1.0.6    2021-08-19 [1] CRAN (R 4.2.0)
 car                    3.0-13   2022-05-02 [1] CRAN (R 4.2.0)
 carData                3.0-5    2022-01-06 [1] CRAN (R 4.2.0)
 checkmate              2.1.0    2022-04-21 [1] CRAN (R 4.2.0)
 chron                  2.3-56   2020-08-18 [1] CRAN (R 4.2.0)
 cli                    3.3.0    2022-04-25 [1] CRAN (R 4.2.0)
 cluster                2.1.3    2022-03-28 [2] CRAN (R 4.2.0)
 codetools              0.2-18   2020-11-04 [2] CRAN (R 4.2.0)
 colorspace             2.0-3    2022-02-21 [1] CRAN (R 4.2.0)
 compareGroups        * 4.5.1    2021-03-29 [1] CRAN (R 4.2.0)
 crayon                 1.5.1    2022-03-26 [1] CRAN (R 4.2.0)
 curl                   4.3.2    2021-06-23 [1] CRAN (R 4.2.0)
 data.table           * 1.14.2   2021-09-27 [1] CRAN (R 4.2.0)
 DBI                    1.1.2    2021-12-20 [1] CRAN (R 4.2.0)
 dbplyr                 2.1.1    2021-04-06 [1] CRAN (R 4.2.0)
 DECIPHER             * 2.24.0   2022-04-26 [1] Bioconductor
 DelayedArray           0.22.0   2022-04-26 [1] Bioconductor
 DiagrammeR           * 1.0.9    2022-03-05 [1] CRAN (R 4.2.0)
 dichromat              2.0-0.1  2022-05-02 [1] CRAN (R 4.2.0)
 digest                 0.6.29   2021-12-01 [1] CRAN (R 4.2.0)
 dplyr                * 1.0.9    2022-04-28 [1] CRAN (R 4.2.0)
 DT                   * 0.23     2022-05-10 [1] CRAN (R 4.2.0)
 ellipsis               0.3.2    2021-04-29 [1] CRAN (R 4.2.0)
 ensembldb              2.20.1   2022-04-29 [1] Bioconductor
 evaluate               0.15     2022-02-18 [1] CRAN (R 4.2.0)
 excelR               * 0.4.0    2020-03-09 [1] CRAN (R 4.2.0)
 fansi                  1.0.3    2022-03-24 [1] CRAN (R 4.2.0)
 fastmap                1.1.0    2021-01-25 [1] CRAN (R 4.2.0)
 fastmatch              1.1-3    2021-07-23 [1] CRAN (R 4.2.0)
 filelock               1.0.2    2018-10-05 [1] CRAN (R 4.2.0)
 flextable              0.7.0    2022-03-06 [1] CRAN (R 4.2.0)
 foreign                0.8-82   2022-01-16 [2] CRAN (R 4.2.0)
 formattable          * 0.2.1    2021-01-07 [1] CRAN (R 4.2.0)
 Formula                1.2-4    2020-10-16 [1] CRAN (R 4.2.0)
 gdtools                0.2.4    2022-02-14 [1] CRAN (R 4.2.0)
 generics               0.1.2    2022-01-31 [1] CRAN (R 4.2.0)
 GenomeInfoDb         * 1.32.1   2022-04-28 [1] Bioconductor
 GenomeInfoDbData       1.2.8    2022-05-16 [1] Bioconductor
 GenomicAlignments      1.32.0   2022-04-26 [1] Bioconductor
 GenomicFeatures        1.48.1   2022-05-15 [1] Bioconductor
 GenomicRanges        * 1.48.0   2022-04-26 [1] Bioconductor
 ggdendro             * 0.1.23   2022-02-16 [1] CRAN (R 4.2.0)
 ggplot2              * 3.3.6    2022-05-03 [1] CRAN (R 4.2.0)
 ggpubr               * 0.4.0    2020-06-27 [1] CRAN (R 4.2.0)
 ggsignif               0.6.3    2021-09-09 [1] CRAN (R 4.2.0)
 glue                   1.6.2    2022-02-24 [1] CRAN (R 4.2.0)
 gmp                    0.6-5    2022-03-17 [1] CRAN (R 4.2.0)
 gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.2.0)
 gtable                 0.3.0    2019-03-25 [1] CRAN (R 4.2.0)
 Gviz                 * 1.40.1   2022-05-03 [1] Bioconductor
 HardyWeinberg          1.7.5    2022-05-07 [1] CRAN (R 4.2.0)
 Hmisc                  4.7-0    2022-04-19 [1] CRAN (R 4.2.0)
 hms                    1.1.1    2021-09-26 [1] CRAN (R 4.2.0)
 htmlTable              2.4.0    2022-01-04 [1] CRAN (R 4.2.0)
 htmltools            * 0.5.2    2021-08-25 [1] CRAN (R 4.2.0)
 htmlwidgets          * 1.5.4    2021-09-08 [1] CRAN (R 4.2.0)
 httpuv                 1.6.5    2022-01-05 [1] CRAN (R 4.2.0)
 httr                   1.4.3    2022-05-04 [1] CRAN (R 4.2.0)
 igraph                 1.3.1    2022-04-20 [1] CRAN (R 4.2.0)
 IRanges              * 2.30.0   2022-04-26 [1] Bioconductor
 jpeg                   0.1-9    2021-07-24 [1] CRAN (R 4.2.0)
 jquerylib              0.1.4    2021-04-26 [1] CRAN (R 4.2.0)
 jsonlite               1.8.0    2022-02-22 [1] CRAN (R 4.2.0)
 kableExtra             1.3.4    2021-02-20 [1] CRAN (R 4.2.0)
 KEGGREST               1.36.0   2022-04-26 [1] Bioconductor
 knitr                * 1.39     2022-04-26 [1] CRAN (R 4.2.0)
 later                  1.3.0    2021-08-18 [1] CRAN (R 4.2.0)
 lattice              * 0.20-45  2021-09-22 [2] CRAN (R 4.2.0)
 latticeExtra           0.6-29   2019-12-19 [1] CRAN (R 4.2.0)
 lazyeval               0.2.2    2019-03-15 [1] CRAN (R 4.2.0)
 lifecycle              1.0.1    2021-09-24 [1] CRAN (R 4.2.0)
 logger               * 0.2.2    2021-10-19 [1] CRAN (R 4.2.0)
 magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.2.0)
 MASS                   7.3-56   2022-03-23 [2] CRAN (R 4.2.0)
 Matrix                 1.4-1    2022-03-23 [2] CRAN (R 4.2.0)
 MatrixGenerics         1.8.0    2022-04-26 [1] Bioconductor
 matrixStats            0.62.0   2022-04-19 [1] CRAN (R 4.2.0)
 memoise                2.0.1    2021-11-26 [1] CRAN (R 4.2.0)
 mice                   3.14.0   2021-11-24 [1] CRAN (R 4.2.0)
 mime                   0.12     2021-09-28 [1] CRAN (R 4.2.0)
 munsell                0.5.0    2018-06-12 [1] CRAN (R 4.2.0)
 nlme                   3.1-157  2022-03-25 [2] CRAN (R 4.2.0)
 nnet                   7.3-17   2022-01-16 [2] CRAN (R 4.2.0)
 officer                0.4.2    2022-03-23 [1] CRAN (R 4.2.0)
 openxlsx             * 4.2.5    2021-12-14 [1] CRAN (R 4.2.0)
 pdftools             * 3.2.0    2022-04-19 [1] CRAN (R 4.2.0)
 phangorn             * 2.8.1    2021-12-15 [1] CRAN (R 4.2.0)
 pillar                 1.7.0    2022-02-01 [1] CRAN (R 4.2.0)
 pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.2.0)
 plotly               * 4.10.0   2021-10-09 [1] CRAN (R 4.2.0)
 plyr                 * 1.8.7    2022-03-24 [1] CRAN (R 4.2.0)
 png                  * 0.1-7    2013-12-03 [1] CRAN (R 4.2.0)
 prettyunits            1.1.1    2020-01-24 [1] CRAN (R 4.2.0)
 progress               1.2.2    2019-05-16 [1] CRAN (R 4.2.0)
 promises               1.2.0.1  2021-02-11 [1] CRAN (R 4.2.0)
 ProtGenerics           1.28.0   2022-04-26 [1] Bioconductor
 purrr                * 0.3.4    2020-04-17 [1] CRAN (R 4.2.0)
 qpdf                   1.1      2019-03-07 [1] CRAN (R 4.2.0)
 quadprog               1.5-8    2019-11-20 [1] CRAN (R 4.2.0)
 R6                     2.5.1    2021-08-19 [1] CRAN (R 4.2.0)
 rappdirs               0.3.3    2021-01-31 [1] CRAN (R 4.2.0)
 RColorBrewer         * 1.1-3    2022-04-03 [1] CRAN (R 4.2.0)
 Rcpp                   1.0.8.3  2022-03-17 [1] CRAN (R 4.2.0)
 RCurl                  1.98-1.6 2022-02-08 [1] CRAN (R 4.2.0)
 readr                * 2.1.2    2022-01-30 [1] CRAN (R 4.2.0)
 renv                 * 0.15.4   2022-03-03 [1] CRAN (R 4.2.0)
 reshape2             * 1.4.4    2020-04-09 [1] CRAN (R 4.2.0)
 restfulr               0.0.13   2017-08-06 [1] CRAN (R 4.2.0)
 rjson                  0.2.21   2022-01-09 [1] CRAN (R 4.2.0)
 rlang                  1.0.2    2022-03-04 [1] CRAN (R 4.2.0)
 rlist                * 0.4.6.2  2021-09-03 [1] CRAN (R 4.2.0)
 rmarkdown            * 2.14     2022-04-25 [1] CRAN (R 4.2.0)
 Rmisc                * 1.5.1    2022-05-02 [1] CRAN (R 4.2.0)
 rpart                  4.1.16   2022-01-24 [2] CRAN (R 4.2.0)
 Rsamtools              2.12.0   2022-04-26 [1] Bioconductor
 Rsolnp                 1.16     2015-12-28 [1] CRAN (R 4.2.0)
 RSQLite              * 2.2.14   2022-05-07 [1] CRAN (R 4.2.0)
 rstatix              * 0.7.0    2021-02-13 [1] CRAN (R 4.2.0)
 rstudioapi             0.13     2020-11-12 [1] CRAN (R 4.2.0)
 rtracklayer          * 1.56.0   2022-04-29 [1] Bioconductor
 rvest                  1.0.2    2021-10-16 [1] CRAN (R 4.2.0)
 S4Vectors            * 0.34.0   2022-04-26 [1] Bioconductor
 sangeranalyseR       * 1.6.1    2022-05-15 [1] Bioconductor
 sangerseqR           * 1.32.0   2022-04-26 [1] Bioconductor
 sass                   0.4.1    2022-03-23 [1] CRAN (R 4.2.0)
 scales                 1.2.0    2022-04-13 [1] CRAN (R 4.2.0)
 seqinr               * 4.2-8    2021-06-09 [1] CRAN (R 4.2.0)
 sessioninfo            1.2.2    2021-12-06 [1] CRAN (R 4.2.0)
 shiny                * 1.7.1    2021-10-02 [1] CRAN (R 4.2.0)
 shinycssloaders      * 1.0.0    2020-07-28 [1] CRAN (R 4.2.0)
 shinydashboard       * 0.7.2    2021-09-30 [1] CRAN (R 4.2.0)
 shinyjs              * 2.1.0    2021-12-23 [1] CRAN (R 4.2.0)
 shinythemes          * 1.2.0    2021-01-25 [1] CRAN (R 4.2.0)
 shinyWidgets         * 0.7.0    2022-05-11 [1] CRAN (R 4.2.0)
 stringi                1.7.6    2021-11-29 [1] CRAN (R 4.2.0)
 stringr              * 1.4.0    2019-02-10 [1] CRAN (R 4.2.0)
 SummarizedExperiment   1.26.1   2022-04-29 [1] Bioconductor
 survival               3.3-1    2022-03-03 [2] CRAN (R 4.2.0)
 svglite                2.1.0    2022-02-03 [1] CRAN (R 4.2.0)
 systemfonts            1.0.4    2022-02-11 [1] CRAN (R 4.2.0)
 tibble                 3.1.7    2022-05-03 [1] CRAN (R 4.2.0)
 tidyr                  1.2.0    2022-02-01 [1] CRAN (R 4.2.0)
 tidyselect             1.1.2    2022-02-21 [1] CRAN (R 4.2.0)
 truncnorm              1.0-8    2018-02-27 [1] CRAN (R 4.2.0)
 tzdb                   0.3.0    2022-03-28 [1] CRAN (R 4.2.0)
 utf8                   1.2.2    2021-07-24 [1] CRAN (R 4.2.0)
 uuid                   1.1-0    2022-04-19 [1] CRAN (R 4.2.0)
 VariantAnnotation      1.42.0   2022-04-26 [1] Bioconductor
 vctrs                  0.4.1    2022-04-13 [1] CRAN (R 4.2.0)
 viridisLite            0.4.0    2021-04-13 [1] CRAN (R 4.2.0)
 visNetwork             2.1.0    2021-09-29 [1] CRAN (R 4.2.0)
 webshot              * 0.5.3    2022-04-14 [1] CRAN (R 4.2.0)
 withr                  2.5.0    2022-03-03 [1] CRAN (R 4.2.0)
 writexl                1.4.0    2021-04-20 [1] CRAN (R 4.2.0)
 xfun                   0.31     2022-05-10 [1] CRAN (R 4.2.0)
 XML                    3.99-0.9 2022-02-24 [1] CRAN (R 4.2.0)
 xml2                   1.3.3    2021-11-30 [1] CRAN (R 4.2.0)
 xtable                 1.8-4    2019-04-21 [1] CRAN (R 4.2.0)
 XVector              * 0.36.0   2022-04-26 [1] Bioconductor
 yaml                   2.3.5    2022-02-21 [1] CRAN (R 4.2.0)
 zeallot              * 0.1.0    2018-01-28 [1] CRAN (R 4.2.0)
 zip                    2.2.0    2021-05-31 [1] CRAN (R 4.2.0)
 zlibbioc               1.42.0   2022-04-26 [1] Bioconductor

 [1] D:/Documents/These/ABSP/renv/library/R-4.2/x86_64-w64-mingw32
 [2] C:/Program Files/R/R-4.2.0/library




> sessionInfo()
R version 4.2.0 (2022-04-22 ucrt)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 19044)

Matrix products: default

locale:
[1] LC_COLLATE=French_France.utf8  LC_CTYPE=French_France.utf8    LC_MONETARY=French_France.utf8 LC_NUMERIC=C                   LC_TIME=French_France.utf8    

attached base packages:
 [1] grDevices datasets  tools     parallel  grid      stats4    stats     graphics  utils     methods   base     

other attached packages:
 [1] webshot_0.5.3         shinythemes_1.2.0     sangeranalyseR_1.6.1  logger_0.2.2          BiocStyle_2.24.0      seqinr_4.2-8          shinyWidgets_0.7.0   
 [8] shinycssloaders_1.0.0 excelR_0.4.0          zeallot_0.1.0         DT_0.23               data.table_1.14.2     shinyjs_2.1.0         shinydashboard_0.7.2 
[15] shiny_1.7.1           gridExtra_2.3         sangerseqR_1.32.0     phangorn_2.8.1        reshape2_1.4.4        DECIPHER_2.24.0       RSQLite_2.2.14       
[22] ape_5.6-2             stringr_1.4.0         rstatix_0.7.0         Rmisc_1.5.1           plyr_1.8.7            lattice_0.20-45       rmarkdown_2.14       
[29] rlist_0.4.6.2         renv_0.15.4           readr_2.1.2           RColorBrewer_1.1-3    purrr_0.3.4           png_0.1-7             plotly_4.10.0        
[36] pdftools_3.2.0        openxlsx_4.2.5        knitr_1.39            htmlwidgets_1.5.4     htmltools_0.5.2       Gviz_1.40.1           ggpubr_0.4.0         
[43] ggplot2_3.3.6         ggdendro_0.1.23       formattable_0.2.1     dplyr_1.0.9           DiagrammeR_1.0.9      compareGroups_4.5.1   BSgenome_1.64.0      
[50] rtracklayer_1.56.0    GenomicRanges_1.48.0  Biostrings_2.64.0     GenomeInfoDb_1.32.1   XVector_0.36.0        IRanges_2.30.0        S4Vectors_0.34.0     
[57] BiocManager_1.30.17   BiocGenerics_0.42.0   arrangements_1.1.9   

loaded via a namespace (and not attached):
  [1] utf8_1.2.2                  tidyselect_1.1.2            AnnotationDbi_1.58.0        BiocParallel_1.30.0         gmp_0.6-5                   munsell_0.5.0              
  [7] codetools_0.2-18            chron_2.3-56                withr_2.5.0                 colorspace_2.0-3            Biobase_2.56.0              filelock_1.0.2             
 [13] uuid_1.1-0                  rstudioapi_0.13             ggsignif_0.6.3              officer_0.4.2               MatrixGenerics_1.8.0        GenomeInfoDbData_1.2.8     
 [19] bit64_4.0.5                 vctrs_0.4.1                 generics_0.1.2              xfun_0.31                   biovizBase_1.44.0           BiocFileCache_2.4.0        
 [25] qpdf_1.1                    R6_2.5.1                    AnnotationFilter_1.20.0     bitops_1.0-7                cachem_1.0.6                DelayedArray_0.22.0        
 [31] assertthat_0.2.1            promises_1.2.0.1            BiocIO_1.6.0                scales_1.2.0                nnet_7.3-17                 gtable_0.3.0               
 [37] ensembldb_2.20.1            rlang_1.0.2                 systemfonts_1.0.4           splines_4.2.0               lazyeval_0.2.2              dichromat_2.0-0.1          
 [43] broom_0.8.0                 checkmate_2.1.0             yaml_2.3.5                  abind_1.4-5                 GenomicFeatures_1.48.1      backports_1.4.1            
 [49] httpuv_1.6.5                HardyWeinberg_1.7.5         Hmisc_4.7-0                 ellipsis_0.3.2              kableExtra_1.3.4            jquerylib_0.1.4            
 [55] Rsolnp_1.16                 Rcpp_1.0.8.3                base64enc_0.1-3             visNetwork_2.1.0            progress_1.2.2              zlibbioc_1.42.0            
 [61] RCurl_1.98-1.6              prettyunits_1.1.1           rpart_4.1.16                SummarizedExperiment_1.26.1 cluster_2.1.3               magrittr_2.0.3             
 [67] flextable_0.7.0             truncnorm_1.0-8             ProtGenerics_1.28.0         matrixStats_0.62.0          xtable_1.8-4                mime_0.12                  
 [73] hms_1.1.1                   evaluate_0.15               XML_3.99-0.9                jpeg_0.1-9                  compiler_4.2.0              biomaRt_2.52.0             
 [79] tibble_3.1.7                mice_3.14.0                 writexl_1.4.0               crayon_1.5.1                later_1.3.0                 tzdb_0.3.0                 
 [85] Formula_1.2-4               tidyr_1.2.0                 DBI_1.1.2                   dbplyr_2.1.1                MASS_7.3-56                 rappdirs_0.3.3             
 [91] ade4_1.7-19                 Matrix_1.4-1                car_3.0-13                  cli_3.3.0                   quadprog_1.5-8              igraph_1.3.1               
 [97] pkgconfig_2.0.3             GenomicAlignments_1.32.0    foreign_0.8-82              xml2_1.3.3                  svglite_2.1.0               bslib_0.3.1                
[103] rvest_1.0.2                 VariantAnnotation_1.42.0    digest_0.6.29               fastmatch_1.1-3             htmlTable_2.4.0             gdtools_0.2.4              
[109] restfulr_0.0.13             curl_4.3.2                  Rsamtools_2.12.0            rjson_0.2.21                lifecycle_1.0.1             nlme_3.1-157               
[115] jsonlite_1.8.0              carData_3.0-5               viridisLite_0.4.0           askpass_1.1                 fansi_1.0.3                 pillar_1.7.0               
[121] KEGGREST_1.36.0             fastmap_1.1.0               httr_1.4.3                  survival_3.3-1              glue_1.6.2                  zip_2.2.0                  
[127] bit_4.0.4                   sass_0.4.1                  stringi_1.7.6               blob_1.2.3                  latticeExtra_0.6-29         memoise_2.0.1  

-->
